### 2012-05-26 ras2012-paper/src/main/latex/paper/paper.tex
Add the paper module.

\documentclass[10pt,a4paper,draft]{article}
\usepackage[utf8x]{inputenc}
\usepackage{ucs}
\usepackage[english]{babel}
\usepackage{multirow}
\author{Lukáš Petrovický}
\title{RAS2012: Competition Entry}
\begin{document}

\maketitle
\listoffigures
\listoftables
\tableofcontents

\section{Hello...}

...world!

\appendix

\input{RDS2.txt-13829.tex}.

\end{document}

### 2012-05-26 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/RAS2012Solution.java
Remove Solution's hashCode/equals so that the solution tabu starts working.

@@ -75,28 +75,6 @@ public class RAS2012Solution extends Visualizable implements Solution<HardAndSof
         return solution;
     }
 
-    @Override
-    public boolean equals(final Object obj) {
-        if (this == obj) {
-            return true;
-        }
-        if (obj == null) {
-            return false;
-        }
-        if (!(obj instanceof RAS2012Solution)) {
-            return false;
-        }
-        final RAS2012Solution other = (RAS2012Solution) obj;
-        if (this.assignments == null) {
-            if (other.assignments != null) {
-                return false;
-            }
-        } else if (!this.assignments.equals(other.assignments)) {
-            return false;
-        }
-        return true;
-    }
-
     public ItineraryAssignment getAssignment(final Train t) {
         return this.assignments.get(t);
     }
@@ -146,14 +124,6 @@ public class RAS2012Solution extends Visualizable implements Solution<HardAndSof
 
     public SortedSet<Train> getTrains() {
         return new TreeSet<Train>(this.trains);
-    }
-
-    @Override
-    public int hashCode() {
-        final int prime = 31;
-        int result = 1;
-        result = prime * result + (this.assignments == null ? 0 : this.assignments.hashCode());
-        return result;
     }
 
     @Override

### 2012-05-26 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/ItineraryAssignment.java
Remove ItineraryAssignment's hashCode/equals so that the entity tabu starts working.

@@ -37,28 +37,6 @@ public final class ItineraryAssignment implements Cloneable {
         return ia;
     }
 
-    @Override
-    public boolean equals(final Object obj) {
-        if (this == obj) {
-            return true;
-        }
-        if (obj == null) {
-            return false;
-        }
-        if (!(obj instanceof ItineraryAssignment)) {
-            return false;
-        }
-        final ItineraryAssignment other = (ItineraryAssignment) obj;
-        if (this.itinerary == null) {
-            if (other.itinerary != null) {
-                return false;
-            }
-        } else if (!this.itinerary.equals(other.itinerary)) {
-            return false;
-        }
-        return true;
-    }
-
     public Itinerary getItinerary() {
         if (this.itinerary == null) {
             throw new IllegalStateException("No itinerary available, provide a route first.");
@@ -74,14 +52,6 @@ public final class ItineraryAssignment implements Cloneable {
 
     public Train getTrain() {
         return this.train;
-    }
-
-    @Override
-    public int hashCode() {
-        final int prime = 31;
-        int result = 1;
-        result = prime * result + (this.itinerary == null ? 0 : this.itinerary.hashCode());
-        return result;
     }
 
     public synchronized void setRoute(final Route route) {

### 2012-05-26 ras2012-solver/src/test/java/org/drools/planner/examples/ras2012/model/ItineraryAssignmentTest.java

@@ -57,7 +57,6 @@ public class ItineraryAssignmentTest extends AbstractItineraryProviderBasedTest 
         ia.setRoute(this.expectedRoute);
         final ItineraryAssignment ia2 = ia.clone();
         Assert.assertNotSame(ia2, ia);
-        Assert.assertEquals(ia2, ia);
         Assert.assertSame(ia.getItinerary(), ia2.getItinerary());
     }
 
@@ -78,31 +77,6 @@ public class ItineraryAssignmentTest extends AbstractItineraryProviderBasedTest 
     @Test(expected = IllegalArgumentException.class)
     public void testConstructorWithNullTrain() {
         new ItineraryAssignment(null, this.solution.getMaintenances());
-    }
-
-    @Test
-    public void testEqualsExisting() {
-        final ItineraryAssignment ia = new ItineraryAssignment(this.expectedTrain,
-                this.solution.getMaintenances());
-        Assert.assertEquals(ia, ia); // equals itself
-        final ItineraryAssignment ia2 = new ItineraryAssignment(this.expectedTrain,
-                this.solution.getMaintenances());
-        Assert.assertEquals(ia2, ia2);
-        Assert.assertEquals(ia, ia2);
-        ia2.setRoute(this.expectedRoute);
-        Assert.assertEquals(ia2, ia2);
-        Assert.assertFalse(ia2.equals(ia));
-        Assert.assertFalse(ia.equals(ia2));
-        ia.setRoute(this.expectedRoute);
-        Assert.assertEquals(ia, ia2);
-    }
-
-    @Test
-    public void testEqualsNonsense() {
-        final ItineraryAssignment ia = new ItineraryAssignment(this.expectedTrain,
-                this.solution.getMaintenances());
-        Assert.assertFalse(ia.equals(null));
-        Assert.assertFalse(ia.equals("Testing string"));
     }
 
     @Test

### 2012-05-26 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/RAS2012ScoreCalculator.java
Remove unnecessary attribute.

@@ -117,7 +117,7 @@ public class RAS2012ScoreCalculator extends AbstractIncrementalScoreCalculator<R
                         * RAS2012ScoreCalculator.OCCUPATION_CHECKS_PER_MINUTE);
     }
 
-    public int getDelayPenalty(final Itinerary i, final RAS2012Solution solution) {
+    public int getDelayPenalty(final Itinerary i) {
         final long delay = i.getDelay();
         if (delay <= 0) {
             return 0;
@@ -196,7 +196,7 @@ public class RAS2012ScoreCalculator extends AbstractIncrementalScoreCalculator<R
         this.unpreferredTracksPenalties.put(t, this.getUnpreferredTracksPenalty(i));
         this.scheduleAdherencePenalties.put(t, this.getScheduleAdherencePenalty(i));
         this.wantTimePenalties.put(t, this.getWantTimePenalty(i));
-        this.delayPenalties.put(t, this.getDelayPenalty(i, this.solution));
+        this.delayPenalties.put(t, this.getDelayPenalty(i));
         this.recalculateOccupiedArcs(ia);
     }
 

### 2012-05-26 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/App.java
Minor updates.

@@ -23,7 +23,7 @@ public class App {
         // read solution
         final File f = new File("src/main/resources/org/drools/planner/examples/ras2012/RDS2.txt");
         final SolutionIO io = new SolutionIO();
-        final RAS2012Solution sol = io.read(f);
+        RAS2012Solution sol = io.read(f);
         // and now start solving
         final XmlSolverFactory configurer = new XmlSolverFactory();
         configurer.configure(App.class.getResourceAsStream("/solverConfig.xml"));
@@ -35,14 +35,12 @@ public class App {
         if (!targetFolder.exists()) {
             targetFolder.mkdirs();
         }
-        final RAS2012Solution solution = (RAS2012Solution) solver.getBestSolution();
-        final HardAndSoftScore score = App.recaculateScore(solution);
+        sol = (RAS2012Solution) solver.getBestSolution();
+        final HardAndSoftScore score = App.recaculateScore(sol);
         if (score.getHardScore() >= 0) { // don't write score that isn't feasible
-            io.writeXML(solution, new File(targetFolder, f.getName() + score.getSoftScore()
-                    + ".xml"));
-            io.writeTex(solution, new File(targetFolder, f.getName() + score.getSoftScore()
-                    + ".tex"));
-            solution.visualize(new File(targetFolder, f.getName() + ".png"));
+            io.writeXML(sol, new File(targetFolder, f.getName() + score.getSoftScore() + ".xml"));
+            io.writeTex(sol, new File(targetFolder, f.getName() + score.getSoftScore() + ".tex"));
+            sol.visualize(new File(targetFolder, f.getName() + score.getSoftScore() + ".png"));
         } else {
             logger.warn("Not writing results because solution wasn't feasible: " + score);
         }

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/ScoreCalculator.java
Only recalculate occupied arcs when actually necessary.

@@ -5,15 +5,19 @@ import java.util.HashMap;
 import java.util.Map;
 import java.util.concurrent.TimeUnit;
 
+import org.apache.commons.lang.NotImplementedException;
+import org.apache.commons.lang3.tuple.Pair;
 import org.drools.planner.core.score.buildin.hardandsoft.DefaultHardAndSoftScore;
 import org.drools.planner.core.score.buildin.hardandsoft.HardAndSoftScore;
 import org.drools.planner.core.score.director.incremental.AbstractIncrementalScoreCalculator;
 import org.drools.planner.examples.ras2012.model.Itinerary;
+import org.drools.planner.examples.ras2012.model.Itinerary.ChangeType;
 import org.drools.planner.examples.ras2012.model.ItineraryAssignment;
 import org.drools.planner.examples.ras2012.model.Node;
 import org.drools.planner.examples.ras2012.model.Train;
 import org.drools.planner.examples.ras2012.util.ConflictRegistry;
 import org.drools.planner.examples.ras2012.util.Converter;
+import org.drools.planner.examples.ras2012.util.model.OccupationTracker;
 import org.slf4j.Logger;
 import org.slf4j.LoggerFactory;
 
@@ -52,53 +56,47 @@ public class ScoreCalculator extends AbstractIncrementalScoreCalculator<ProblemS
     @Override
     public void afterAllVariablesChanged(final Object entity) {
         if (entity instanceof ItineraryAssignment) {
-            this.insert((ItineraryAssignment) entity);
+            this.modify((ItineraryAssignment) entity);
         }
     }
 
     @Override
     public void afterEntityAdded(final Object entity) {
         if (entity instanceof ItineraryAssignment) {
-            this.insert((ItineraryAssignment) entity);
+            this.modify((ItineraryAssignment) entity);
         }
     }
 
     @Override
     public void afterEntityRemoved(final Object entity) {
-        // do nothing
+        throw new NotImplementedException();
     }
 
     @Override
     public void afterVariableChanged(final Object entity, final String variableName) {
         if (entity instanceof ItineraryAssignment) {
-            this.insert((ItineraryAssignment) entity);
+            this.modify((ItineraryAssignment) entity);
         }
     }
 
     @Override
     public void beforeAllVariablesChanged(final Object entity) {
-        if (entity instanceof ItineraryAssignment) {
-            this.retract((ItineraryAssignment) entity);
-        }
+        throw new NotImplementedException();
[…]

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/Itinerary.java

@@ -13,6 +13,8 @@ import java.util.concurrent.atomic.AtomicBoolean;
 
 import org.apache.commons.lang3.builder.EqualsBuilder;
 import org.apache.commons.lang3.builder.HashCodeBuilder;
+import org.apache.commons.lang3.tuple.ImmutablePair;
+import org.apache.commons.lang3.tuple.Pair;
 import org.drools.planner.examples.ras2012.Visualizable;
 import org.drools.planner.examples.ras2012.util.Converter;
 import org.drools.planner.examples.ras2012.util.model.ArcProgression;
@@ -22,6 +24,16 @@ import org.slf4j.Logger;
 import org.slf4j.LoggerFactory;
 
 public final class Itinerary extends Visualizable {
+
+    public static enum ChangeType {
+
+        UNCHANGED, REMOVE_ALL_WAIT_TIMES, REMOVE_WAIT_TIME, SET_WAIT_TIME;
+
+    }
+
+    Pair<ChangeType, Node>                     lastChange            = ImmutablePair.of(
+                                                                             ChangeType.UNCHANGED,
+                                                                             null);
 
     private static final TimeUnit              DEFAULT_TIME_UNIT     = TimeUnit.MILLISECONDS;
 
@@ -177,6 +189,10 @@ public final class Itinerary extends Visualizable {
         return this.delay;
     }
 
+    public Pair<ChangeType, Node> getLatestWaitTimeChange() {
+        return this.lastChange;
+    }
+
     protected Arc getLeadingArc(final long time) {
         if (time < this.trainEntryTime) {
             return null;
@@ -326,6 +342,7 @@ public final class Itinerary extends Visualizable {
         if (this.nodeWaitTimes.containsKey(n)) {
             Itinerary.logger.debug("Removing wait time for {} from {}.", new Object[] { n, this });
             this.invalidateCaches();
+            this.lastChange = ImmutablePair.of(ChangeType.REMOVE_WAIT_TIME, n);
             return this.nodeWaitTimes.remove(n);
         } else {
             Itinerary.logger.debug("No wait time to remove for {} from {}.",
@@ -340,6 +357,7 @@ public final class Itinerary extends Visualizable {
             this.invalidateCaches();
         }
         this.nodeWaitTimes.clear();
+        this.lastChange = ImmutablePair.of(ChangeType.REMOVE_ALL_WAIT_TIMES, null);
     }
 
[…]

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/move/RouteReassignmentMove.java

@@ -44,9 +44,8 @@ public class RouteReassignmentMove implements Move {
     @Override
     public void doMove(final ScoreDirector scoreDirector) {
         this.assignment = this.initializeMove(scoreDirector);
-        scoreDirector.beforeVariableChanged(this.assignment, "route");
         this.assignment.setRoute(this.route);
-        scoreDirector.afterVariableChanged(this.assignment, "route");
+        scoreDirector.afterEntityAdded(this.assignment);
     }
 
     @Override

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/move/RouteReassignmentUndoMove.java

@@ -39,12 +39,11 @@ public class RouteReassignmentUndoMove implements Move {
     @Override
     public void doMove(final ScoreDirector scoreDirector) {
         this.initializeMove(scoreDirector);
-        scoreDirector.beforeVariableChanged(this.assignment, "route");
         this.assignment.setRoute(this.originalRoute);
         for (final Map.Entry<Node, WaitTime> entry : this.originalWaitTimes.entrySet()) {
             this.assignment.getItinerary().setWaitTime(entry.getKey(), entry.getValue());
         }
-        scoreDirector.afterVariableChanged(this.assignment, "route");
+        scoreDirector.afterEntityAdded(this.assignment);
     }
 
     @Override

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/move/WaitTimeAssignmentMove.java

@@ -60,7 +60,6 @@ public class WaitTimeAssignmentMove implements Move {
     public void doMove(final ScoreDirector scoreDirector) {
         this.initializeMove(scoreDirector);
         final ItineraryAssignment ia = this.getAssignment(scoreDirector);
-        scoreDirector.beforeVariableChanged(ia, "waitTime");
         ia.getItinerary().setWaitTime(this.node, this.waitTime);
         scoreDirector.afterVariableChanged(ia, "waitTime");
     }

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/Itinerary.java
Don't synchronize since we only always use it from one thread.

@@ -270,14 +270,14 @@ public final class Itinerary extends Visualizable {
         return this.route;
     }
 
-    public synchronized SortedMap<Long, Node> getSchedule() {
+    public SortedMap<Long, Node> getSchedule() {
         if (!this.scheduleCacheValid.get() || this.scheduleCache.size() == 0) {
             this.cacheSchedule();
         }
         return Collections.unmodifiableSortedMap(this.scheduleCache);
     }
 
-    public synchronized SortedMap<Long, Arc> getScheduleWithArcs() {
+    public SortedMap<Long, Arc> getScheduleWithArcs() {
         if (!this.scheduleCacheValid.get() || this.scheduleCache.size() == 0) {
             this.cacheSchedule();
         }
@@ -315,7 +315,7 @@ public final class Itinerary extends Visualizable {
         return this.train;
     }
 
-    public synchronized WaitTime getWaitTime(final Node n) {
+    public WaitTime getWaitTime(final Node n) {
         return this.nodeWaitTimes.get(n);
     }
 
@@ -329,7 +329,7 @@ public final class Itinerary extends Visualizable {
                 .append(this.nodeWaitTimes).build();
     }
 
-    private synchronized void invalidateCaches() {
+    private void invalidateCaches() {
         this.scheduleCache.clear();
         this.scheduleCacheWithArcs.clear();
     }
@@ -360,7 +360,7 @@ public final class Itinerary extends Visualizable {
         this.lastChange = ImmutablePair.of(ChangeType.REMOVE_ALL_WAIT_TIMES, null);
     }
 
-    public synchronized WaitTime setWaitTime(final Node n, final WaitTime w) {
+    public WaitTime setWaitTime(final Node n, final WaitTime w) {
         if (!this.getRoute().getProgression().getWaitPoints().contains(n)) {
             throw new IllegalArgumentException(n + " not a wait point: " + this);
         }

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/Route.java
Remove unused attribute.

@@ -56,8 +56,8 @@ public class Route extends Visualizable implements Comparable<Route>, Directed {
         }
 
         public Route build() {
-            return new Route(this.idGenerator.getAndAdd(2), this.isEastbound,
-                    this.arcs.toArray(new Arc[this.arcs.size()]));
+            return new Route(this.idGenerator.getAndAdd(2), this.arcs.toArray(new Arc[this.arcs
+                    .size()]));
         }
 
         @Override
@@ -115,7 +115,7 @@ public class Route extends Visualizable implements Comparable<Route>, Directed {
 
     private final Map<Train, Boolean> routePossibilitiesCache      = new HashMap<Train, Boolean>();
 
-    private Route(final int id, final boolean isEastbound, final Arc... e) {
+    private Route(final int id, final Arc... e) {
         this.id = id;
         this.progression = new ArcProgression(this, e);
     }

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/move/WaitTimeAssignmentMove.java
Fix bug with unused attribute.

@@ -40,7 +40,7 @@ public class WaitTimeAssignmentMove implements Move {
         this.route = r;
         this.node = n;
         this.waitTime = wt;
-        this.previousWaitTime = wt;
+        this.previousWaitTime = previousWaitTime;
     }
 
     private boolean assignmentExists(final ScoreDirector scoreDirector) {

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/Itinerary.java
Fix possible incremental calculation bug when using the RouteReassignmentUndoMove.

@@ -193,6 +193,10 @@ public final class Itinerary extends Visualizable {
         return this.lastChange;
     }
 
+    public void resetLatestWaitTimeChange() {
+        this.lastChange = ImmutablePair.of(ChangeType.UNCHANGED, null);
+    }
+
     protected Arc getLeadingArc(final long time) {
         if (time < this.trainEntryTime) {
             return null;

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/move/RouteReassignmentUndoMove.java

@@ -9,6 +9,7 @@ import org.apache.commons.lang3.builder.HashCodeBuilder;
 import org.drools.planner.core.move.Move;
 import org.drools.planner.core.score.director.ScoreDirector;
 import org.drools.planner.examples.ras2012.ProblemSolution;
+import org.drools.planner.examples.ras2012.model.Itinerary;
 import org.drools.planner.examples.ras2012.model.ItineraryAssignment;
 import org.drools.planner.examples.ras2012.model.Node;
 import org.drools.planner.examples.ras2012.model.Route;
@@ -40,9 +41,11 @@ public class RouteReassignmentUndoMove implements Move {
     public void doMove(final ScoreDirector scoreDirector) {
         this.initializeMove(scoreDirector);
         this.assignment.setRoute(this.originalRoute);
+        final Itinerary i = this.assignment.getItinerary();
         for (final Map.Entry<Node, WaitTime> entry : this.originalWaitTimes.entrySet()) {
-            this.assignment.getItinerary().setWaitTime(entry.getKey(), entry.getValue());
+            i.setWaitTime(entry.getKey(), entry.getValue());
         }
+        i.resetLatestWaitTimeChange(); // make sure calculator isn't mistaken the setWaitTime().
         scoreDirector.afterEntityAdded(this.assignment);
     }
 

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/util/ConflictRegistry.java
Cache conflicts to improve performance.

@@ -14,8 +14,13 @@ public class ConflictRegistry {
 
         private final Map<Train, OccupationTracker> occupiedArcsByTrain = new HashMap<Train, OccupationTracker>();
         private final List<OccupationTracker>       occupiedArcs        = new ArrayList<OccupationTracker>();
+        private double                              cache               = Double.MAX_VALUE;
+        private boolean                             isCacheValid        = false;
 
         public double getConflicts() {
+            if (this.isCacheValid) {
+                return this.cache;
+            }
             double conflicts = 0;
             final int size = this.occupiedArcs.size();
             for (int position = 0; position < size; position++) {
@@ -25,22 +30,45 @@ public class ConflictRegistry {
                     conflicts += left.getConflictingMileage(right);
                 }
             }
-            return conflicts;
+            this.cache = conflicts;
+            this.isCacheValid = true;
+            return this.cache;
         }
 
-        public void resetOccupiedArcs(final Train t) {
+        /**
+         * Mark the train as occupying no arcs.
+         * 
+         * @param t
+         * @return True if this change altered the state of the item.
+         */
+        public boolean resetOccupiedArcs(final Train t) {
             final OccupationTracker toRemove = this.occupiedArcsByTrain.remove(t);
             if (toRemove != null) {
                 this.occupiedArcs.remove(toRemove);
+                this.isCacheValid = false;
+                return true;
             }
+            return false;
         }
 
-        public void setOccupiedArcs(final Train t, final OccupationTracker arcs) {
+        /**
+         * Set arcs occupied by the given train.
+         * 
+         * @param t
+         * @param arcs
+         * @return True if this change altered the state of the item.
+         */
+        public boolean setOccupiedArcs(final Train t, final OccupationTracker arcs) {
+            if (arcs.equals(this.occupiedArcsByTrain.get(t))) {
+                return false;
+            }
+            this.isCacheValid = false;
             final OccupationTracker previous = this.occupiedArcsByTrain.put(t, arcs);
             if (previous != null) {
                 this.occupiedArcs.remove(previous);
             }
             this.occupiedArcs.add(arcs);
+            return true;
         }
 
     }

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/ScoreCalculator.java
Improve logging.

@@ -31,7 +31,7 @@ public class ScoreCalculator extends AbstractIncrementalScoreCalculator<ProblemS
     private static final BigDecimal MILLIS_TO_HOURS              = BigDecimal.valueOf(3600000);
 
     public static HardAndSoftScore oneTimeCalculation(final ProblemSolution solution) {
-        final ScoreCalculator calc = new ScoreCalculator();
+        final ScoreCalculator calc = new ScoreCalculator(true);
         calc.resetWorkingSolution(solution);
         return calc.calculateScore();
     }
@@ -40,6 +40,8 @@ public class ScoreCalculator extends AbstractIncrementalScoreCalculator<ProblemS
         return BigDecimal.valueOf(milliseconds).divide(ScoreCalculator.MILLIS_TO_HOURS,
                 Converter.BIGDECIMAL_SCALE, Converter.BIGDECIMAL_ROUNDING);
     }
+
+    private final boolean             talk;
 
     private ProblemSolution           solution                   = null;
 
@@ -52,6 +54,14 @@ public class ScoreCalculator extends AbstractIncrementalScoreCalculator<ProblemS
     private final Map<Train, Integer> unpreferredTracksPenalties = new HashMap<Train, Integer>();
 
     private ConflictRegistry          conflicts;
+
+    public ScoreCalculator() {
+        this(false);
+    }
+
+    private ScoreCalculator(final boolean silent) {
+        this.talk = !silent;
+    }
 
     @Override
     public void afterAllVariablesChanged(final Object entity) {
@@ -101,7 +111,9 @@ public class ScoreCalculator extends AbstractIncrementalScoreCalculator<ProblemS
 
     @Override
     public HardAndSoftScore calculateScore() {
-        ScoreCalculator.logger.debug("Calculating score.");
+        if (this.talk) {
+            ScoreCalculator.logger.debug("Calculating score.");
+        }
         int penalty = 0;
         for (final Train t : this.solution.getTrains()) {
             penalty += this.wantTimePenalties.get(t);
@@ -144,12 +156,17 @@ public class ScoreCalculator extends AbstractIncrementalScoreCalculator<ProblemS
             case REMOVE_WAIT_TIME:
             case SET_WAIT_TIME:
                 // start re-calculating occupied arcs from the first change in the itinerary
-                ScoreCalculator.logger.debug("Last wait time change registered on {} ({}).",
-                        new Object[] { lastChange.getRight(), lastChange.getLeft() });
+                if (this.talk) {
[…]

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/ScoreCalculator.java
Fix potential incremental score calculation bug.

@@ -301,6 +301,7 @@ public class ScoreCalculator extends AbstractIncrementalScoreCalculator<ProblemS
         this.solution = workingSolution;
         this.clearEveryCache();
         for (final ItineraryAssignment ia : this.solution.getAssignments()) {
+            ia.getItinerary().resetLatestWaitTimeChange();
             this.modify(ia);
         }
     }

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/Itinerary.java
Improved logging.

@@ -374,7 +374,7 @@ public final class Itinerary extends Visualizable {
         this.invalidateCaches();
         final WaitTime previous = this.nodeWaitTimes.put(n, w);
         this.lastChange = ImmutablePair.of(ChangeType.SET_WAIT_TIME, n);
-        Itinerary.logger.debug("Set wait time on {} in {}, replacing {}.", new Object[] { n, this,
+        Itinerary.logger.debug("Set {} on {} in {}, replacing {}.", new Object[] { w, n, this,
                 previous });
         return previous;
     }

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/ScoreCalculator.java
Fix score corruption issue.

@@ -155,12 +155,22 @@ public class ScoreCalculator extends AbstractIncrementalScoreCalculator<ProblemS
         switch (lastChange.getLeft()) {
             case REMOVE_WAIT_TIME:
             case SET_WAIT_TIME:
-                // start re-calculating occupied arcs from the first change in the itinerary
-                if (this.talk) {
-                    ScoreCalculator.logger.debug("Last wait time change registered on {} ({}).",
-                            new Object[] { lastChange.getRight(), lastChange.getLeft() });
+                final Node previousToModifiedNode = i.getRoute().getProgression()
+                        .getPrevious(lastChange.getRight());
+                if (previousToModifiedNode != null) {
+                    // start re-calculating occupied arcs from the first change in the itinerary
+                    if (this.talk) {
+                        ScoreCalculator.logger.debug(
+                                "Last wait time change registered on {} ({}).", new Object[] {
+                                        lastChange.getRight(), lastChange.getLeft() });
+                    }
+                    return i.getArrivalTime(previousToModifiedNode);
+                } else {
+                    if (this.talk) {
+                        ScoreCalculator.logger
+                                .debug("Skipping last wait time change, since it is starting the route.");
+                    }
                 }
-                return i.getArrivalTime(lastChange.getRight());
             default:
                 // re-calculate arcs all across the timeline
                 if (this.talk) {

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/Itinerary.java

@@ -167,6 +167,9 @@ public final class Itinerary extends Visualizable {
     }
 
     public long getArrivalTime(final Node n) {
+        if (n == null) {
+            throw new IllegalArgumentException("Node cannot be null.");
+        }
         final SortedMap<Long, Node> nodeEntryTimes = this.getSchedule();
         long timeEntered = -1;
         for (final SortedMap.Entry<Long, Node> entry : nodeEntryTimes.entrySet()) {
@@ -220,6 +223,9 @@ public final class Itinerary extends Visualizable {
     }
 
     public long getLeaveTime(final Node n) {
+        if (n == null) {
+            throw new IllegalArgumentException("Node cannot be null.");
+        }
         final long arrival = this.getArrivalTime(n);
         final Arc a = this.getRoute().getProgression().getWithDestinationNode(n);
         final long travel = Converter.getTimeFromSpeedAndDistance(

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/util/model/ArcProgression.java

@@ -244,6 +244,38 @@ public class ArcProgression implements Directed {
         return this.orderedArcs.peekFirst();
     }
 
+    public Node getPrevious(final Node n) {
+        if (this.nodes.size() == 0) {
+            throw new IllegalArgumentException("No previous node on an empty route.");
+        } else if (n == null) {
+            return this.getDestination().getDestination(this);
+        }
+        final int indexOf = this.nodes.indexOf(n);
+        if (indexOf < 0) {
+            throw new IllegalArgumentException("Node not in the progression!");
+        } else if (indexOf == 0) {
+            return null;
+        } else {
+            return this.nodes.get(indexOf - 1);
+        }
+    }
+
+    public Node getNext(final Node n) {
+        if (this.nodes.size() == 0) {
+            throw new IllegalArgumentException("No next node on an empty route.");
+        } else if (n == null) {
+            return this.getOrigin().getOrigin(this);
+        }
+        final int indexOf = this.nodes.indexOf(n);
+        if (indexOf < 0) {
+            throw new IllegalArgumentException("Node not in the progression!");
+        } else if (indexOf == (this.nodes.size() - 1)) {
+            return null;
+        } else {
+            return this.nodes.get(indexOf + 1);
+        }
+    }
+
     public Arc getPrevious(final Arc a) {
         if (this.orderedArcs.size() == 0) {
             throw new IllegalArgumentException("No previous arc on an empty route.");
@@ -258,7 +290,6 @@ public class ArcProgression implements Directed {
         } else {
             return this.orderedArcs.get(indexOf - 1);
         }
-
     }
 
     private Node getStartingNode(final Directed r, final Collection<Arc> arcs) {

### 2012-06-02 ras2012-solver/src/test/java/org/drools/planner/examples/ras2012/util/model/ArcProgressionTest.java

@@ -180,16 +180,16 @@ public class ArcProgressionTest {
         final Builder b = this.builder.add(a1);
         Route r = b.build();
         Assert.assertSame("On a route with single arc, null next arc is the first one.", r
-                .getProgression().getOrigin(), r.getProgression().getNext(null));
+                .getProgression().getOrigin(), r.getProgression().getNext((Arc) null));
         r = b.add(a2).build();
         Assert.assertSame("On a route with two arcs, null next arc is still the first one.", r
-                .getProgression().getOrigin(), r.getProgression().getNext(null));
+                .getProgression().getOrigin(), r.getProgression().getNext((Arc) null));
     }
 
     @Test(expected = IllegalArgumentException.class)
     public void testGetNextNullEmptyRoute() {
         final Route r = this.route;
-        r.getProgression().getNext(null);
+        r.getProgression().getNext((Arc) null);
     }
 
     @Test
@@ -253,16 +253,16 @@ public class ArcProgressionTest {
         final Builder b = this.builder.add(a1);
         Route r = b.build();
         Assert.assertSame("On a route with single arc, null previous arc is the first one.", r
-                .getProgression().getDestination(), r.getProgression().getPrevious(null));
+                .getProgression().getDestination(), r.getProgression().getPrevious((Arc) null));
         r = b.add(a2).build();
         Assert.assertSame("On a route with two arcs, null previous arc is the last one.", r
-                .getProgression().getDestination(), r.getProgression().getPrevious(null));
+                .getProgression().getDestination(), r.getProgression().getPrevious((Arc) null));
     }
 
     @Test(expected = IllegalArgumentException.class)
     public void testGetPreviousNullEmptyRoute() {
         final Route r = this.route;
-        r.getProgression().getPrevious(null);
+        r.getProgression().getPrevious((Arc) null);
     }
 
     @Test

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/Arc.java
Arc.equals() is called A LOT, optimize performance.

@@ -3,7 +3,6 @@ package org.drools.planner.examples.ras2012.model;
 import java.math.BigDecimal;
 import java.util.concurrent.atomic.AtomicInteger;
 
-import org.apache.commons.lang3.builder.EqualsBuilder;
 import org.apache.commons.lang3.builder.HashCodeBuilder;
 
 public class Arc extends Section {
@@ -44,7 +43,7 @@ public class Arc extends Section {
             return false;
         }
         final Arc rhs = (Arc) obj;
-        return new EqualsBuilder().append(this.id, rhs.id).isEquals();
+        return this.id == rhs.id;
     }
 
     public BigDecimal getLength() {

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/util/model/ArcProgression.java
Cache ArcProgression length.

@@ -31,6 +31,7 @@ public class ArcProgression implements Directed {
     private final List<Node>                 nodes              = new ArrayList<Node>();
     private final Collection<Node>           waitPoints;
     private final boolean                    isEastbound;
+    private final BigDecimal                 length;
 
     public ArcProgression(final Directed directed, final Arc... arcs) {
         this(directed, Arrays.asList(arcs));
@@ -58,6 +59,7 @@ public class ArcProgression implements Directed {
             this.nodes.add(a.getOrigin(this));
             milestone = milestone.add(a.getLength());
         }
+        this.length = milestone;
         if (this.orderedArcs.size() > 0) {
             this.nodes.add(this.getDestination().getDestination(this));
         }
@@ -170,11 +172,7 @@ public class ArcProgression implements Directed {
     }
 
     public BigDecimal getLength() {
-        BigDecimal length = BigDecimal.ZERO;
-        for (final Arc a : this.orderedArcs) {
-            length = length.add(a.getLength());
-        }
-        return length;
+        return this.length;
     }
 
     public Arc getNext(final Arc a) {

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/util/model/ArcProgression.java
Hugely improve the performance of ArcProgression.getDistance()

@@ -135,11 +135,6 @@ public class ArcProgression implements Directed {
     }
 
     public BigDecimal getDistance(final Node end) {
-        // first solve some corner cases
-        if (!this.nodes.contains(end)) {
-            throw new IllegalArgumentException(end + " not in progression!");
-        }
-        // then make sure nodes are in a proper order
         final Node start = this.getOrigin().getOrigin(this);
         return this.getDistance(start, end);
     }
@@ -156,17 +151,13 @@ public class ArcProgression implements Directed {
         // then make sure nodes are in a proper order
         final int startIndex = this.nodes.indexOf(start);
         final int endIndex = this.nodes.indexOf(end);
-        if (startIndex > endIndex) {
-            return this.getDistance(end, start);
-        }
         // and then retrieve the actual distance
+        final int startWith = Math.min(startIndex, endIndex);
+        final int endWith = Math.max(startIndex, endIndex);
         BigDecimal result = BigDecimal.ZERO;
-        for (final Node n : this.nodes.subList(startIndex, endIndex + 1)) {
-            final Arc a = this.arcsPerOrigin.get(n);
+        for (int i = startWith; i <= endWith; i++) {
+            final Arc a = this.getWithOriginNode(this.nodes.get(i));
             result = result.add(a.getLength());
-            if (a.getDestination(this) == end) {
-                break;
-            }
         }
         return result;
     }

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/util/model/ArcProgression.java
Code cleanup.

@@ -183,6 +183,22 @@ public class ArcProgression implements Directed {
 
     }
 
+    public Node getNext(final Node n) {
+        if (this.nodes.size() == 0) {
+            throw new IllegalArgumentException("No next node on an empty route.");
+        } else if (n == null) {
+            return this.getOrigin().getOrigin(this);
+        }
+        final int indexOf = this.nodes.indexOf(n);
+        if (indexOf < 0) {
+            throw new IllegalArgumentException("Node not in the progression!");
+        } else if (indexOf == this.nodes.size() - 1) {
+            return null;
+        } else {
+            return this.nodes.get(indexOf + 1);
+        }
+    }
+
     public List<Node> getNodes() {
         return Collections.unmodifiableList(this.nodes);
     }
@@ -233,38 +249,6 @@ public class ArcProgression implements Directed {
         return this.orderedArcs.peekFirst();
     }
 
-    public Node getPrevious(final Node n) {
-        if (this.nodes.size() == 0) {
-            throw new IllegalArgumentException("No previous node on an empty route.");
-        } else if (n == null) {
-            return this.getDestination().getDestination(this);
-        }
-        final int indexOf = this.nodes.indexOf(n);
-        if (indexOf < 0) {
-            throw new IllegalArgumentException("Node not in the progression!");
-        } else if (indexOf == 0) {
-            return null;
-        } else {
-            return this.nodes.get(indexOf - 1);
-        }
-    }
-
-    public Node getNext(final Node n) {
-        if (this.nodes.size() == 0) {
-            throw new IllegalArgumentException("No next node on an empty route.");
-        } else if (n == null) {
-            return this.getOrigin().getOrigin(this);
-        }
-        final int indexOf = this.nodes.indexOf(n);
-        if (indexOf < 0) {
-            throw new IllegalArgumentException("Node not in the progression!");
-        } else if (indexOf == (this.nodes.size() - 1)) {
-            return null;
-        } else {
-            return this.nodes.get(indexOf + 1);
-        }
-    }
-
     public Arc getPrevious(final Arc a) {
         if (this.orderedArcs.size() == 0) {
             throw new IllegalArgumentException("No previous arc on an empty route.");
@@ -278,6 +262,22 @@ public class ArcProgression implements Directed {
             return null;
         } else {
             return this.orderedArcs.get(indexOf - 1);
+        }
+    }
+
+    public Node getPrevious(final Node n) {
+        if (this.nodes.size() == 0) {
[…]

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/util/model/ArcProgression.java
Fix a bug introduced in the second-to-last commit.

@@ -155,7 +155,7 @@ public class ArcProgression implements Directed {
         final int startWith = Math.min(startIndex, endIndex);
         final int endWith = Math.max(startIndex, endIndex);
         BigDecimal result = BigDecimal.ZERO;
-        for (int i = startWith; i <= endWith; i++) {
+        for (int i = startWith; i < endWith; i++) {
             final Arc a = this.getWithOriginNode(this.nodes.get(i));
             result = result.add(a.getLength());
         }

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/util/model/ArcProgression.java
No need to query for the collection size all the time.

@@ -31,6 +31,7 @@ public class ArcProgression implements Directed {
     private final List<Node>                 nodes              = new ArrayList<Node>();
     private final Collection<Node>           waitPoints;
     private final boolean                    isEastbound;
+    private final boolean                    isEmpty;
     private final BigDecimal                 length;
 
     public ArcProgression(final Directed directed, final Arc... arcs) {
@@ -60,7 +61,8 @@ public class ArcProgression implements Directed {
             milestone = milestone.add(a.getLength());
         }
         this.length = milestone;
-        if (this.orderedArcs.size() > 0) {
+        this.isEmpty = this.orderedArcs.size() == 0;
+        if (!this.isEmpty) {
             this.nodes.add(this.getDestination().getDestination(this));
         }
         // determine whether a particular arc is preferred
@@ -73,7 +75,7 @@ public class ArcProgression implements Directed {
 
     private Collection<Node> assembleWaitPoints() {
         final Collection<Node> points = new TreeSet<Node>();
-        if (this.orderedArcs.size() == 0) {
+        if (this.isEmpty) {
             return Collections.unmodifiableCollection(points);
         }
         // we want to be able to hold the train before it enters the network
@@ -167,7 +169,7 @@ public class ArcProgression implements Directed {
     }
 
     public Arc getNext(final Arc a) {
-        if (this.orderedArcs.size() == 0) {
+        if (this.isEmpty) {
             throw new IllegalArgumentException("No next arc on an empty route.");
         } else if (a == null) {
             return this.getOrigin();
@@ -184,7 +186,7 @@ public class ArcProgression implements Directed {
     }
 
     public Node getNext(final Node n) {
-        if (this.nodes.size() == 0) {
+        if (this.isEmpty) {
             throw new IllegalArgumentException("No next node on an empty route.");
         } else if (n == null) {
             return this.getOrigin().getOrigin(this);
@@ -250,7 +252,7 @@ public class ArcProgression implements Directed {
     }
 
     public Arc getPrevious(final Arc a) {
-        if (this.orderedArcs.size() == 0) {
+        if (this.isEmpty) {
             throw new IllegalArgumentException("No previous arc on an empty route.");
         } else if (a == null) {
             return this.getDestination();
@@ -266,7 +268,7 @@ public class ArcProgression implements Directed {
     }
 
     public Node getPrevious(final Node n) {
-        if (this.nodes.size() == 0) {
+        if (this.isEmpty) {
[…]

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/util/model/ArcProgression.java
Don't create collection unless necessary.

@@ -73,11 +73,12 @@ public class ArcProgression implements Directed {
         this.waitPoints = this.assembleWaitPoints();
     }
 
+    @SuppressWarnings("unchecked")
     private Collection<Node> assembleWaitPoints() {
-        final Collection<Node> points = new TreeSet<Node>();
         if (this.isEmpty) {
-            return Collections.unmodifiableCollection(points);
+            return Collections.unmodifiableSet(Collections.EMPTY_SET);
         }
+        final Collection<Node> points = new TreeSet<Node>();
         // we want to be able to hold the train before it enters the network
         final Arc firstArc = this.getOrigin();
         points.add(firstArc.getOrigin(this));

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/util/model/ArcProgression.java
Further optimize getDistance().

@@ -143,20 +143,22 @@ public class ArcProgression implements Directed {
     }
 
     public BigDecimal getDistance(final Node start, final Node end) {
-        // first solve some corner cases
-        if (!this.nodes.contains(start)) {
-            throw new IllegalArgumentException(start + " not in progression!");
-        } else if (!this.nodes.contains(end)) {
-            throw new IllegalArgumentException(end + " not in progression!");
-        } else if (start == end) {
+        if (start == end) {
             return BigDecimal.ZERO;
         }
-        // then make sure nodes are in a proper order
+        // locate nodes in the collection
         final int startIndex = this.nodes.indexOf(start);
+        if (startIndex < 0) {
+            throw new IllegalArgumentException(start + " not in progression.");
+        }
         final int endIndex = this.nodes.indexOf(end);
-        // and then retrieve the actual distance
+        if (endIndex < 0) {
+            throw new IllegalArgumentException(end + " not in progression.");
+        }
+        // make sure they are in a proper order
         final int startWith = Math.min(startIndex, endIndex);
         final int endWith = Math.max(startIndex, endIndex);
+        // and then retrieve the actual distance
         BigDecimal result = BigDecimal.ZERO;
         for (int i = startWith; i < endWith; i++) {
             final Arc a = this.getWithOriginNode(this.nodes.get(i));

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/Itinerary.java
Performance improvements to Itinerary.getOccupiedArcs()

@@ -44,6 +44,7 @@ public final class Itinerary extends Visualizable {
 
     private final AtomicBoolean                scheduleCacheValid    = new AtomicBoolean(false);
     private final Collection<Node>             nodesEnRoute;
+    private final Collection<Arc>              arcsEnRoute;
     private final SortedMap<Long, Node>        scheduleCache         = new TreeMap<Long, Node>();
     private final SortedMap<Long, Arc>         scheduleCacheWithArcs = new TreeMap<Long, Arc>();
     private final long                         trainEntryTime;
@@ -71,8 +72,10 @@ public final class Itinerary extends Visualizable {
         this.route = r;
         this.train = t;
         this.trainEntryTime = t.getEntryTime(Itinerary.DEFAULT_TIME_UNIT);
-        this.nodesEnRoute = this.getRoute().getProgression().tail(this.getTrain().getOrigin())
-                .getNodes();
+        final ArcProgression enRoute = this.getRoute().getProgression()
+                .tail(this.getTrain().getOrigin());
+        this.nodesEnRoute = enRoute.getNodes();
+        this.arcsEnRoute = enRoute.getArcs();
         // initialize the maintenance windows
         if (maintenanceWindows != null) {
             for (final MaintenanceWindow mow : maintenanceWindows) {
@@ -90,16 +93,7 @@ public final class Itinerary extends Visualizable {
         int i = 0;
         long previousTime = 0;
         Arc previousArc = null;
-        boolean seekingStart = true;
-        for (final Arc currentArc : this.getRoute().getProgression().getArcs()) {
-            // arc progression begins with the start of the route; the train doesn't necessarily start there
-            if (seekingStart) {
-                if (currentArc.getOrigin(this.getTrain()) != this.getTrain().getOrigin()) {
-                    continue;
-                } else {
-                    seekingStart = false;
-                }
-            }
+        for (final Arc currentArc : this.arcsEnRoute) {
             long time = 0;
             if (i == 0) {
                 // first item needs to be augmented by the train entry time
@@ -196,10 +190,6 @@ public final class Itinerary extends Visualizable {
         return this.lastChange;
     }
 
-    public void resetLatestWaitTimeChange() {
-        this.lastChange = ImmutablePair.of(ChangeType.UNCHANGED, null);
-    }
-
     protected Arc getLeadingArc(final long time) {
         if (time < this.trainEntryTime) {
             return null;
@@ -207,10 +197,13 @@ public final class Itinerary extends Visualizable {
[…]

### 2012-06-02 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/util/model/ArcProgression.java
Better collection.

@@ -7,7 +7,6 @@ import java.util.Collection;
 import java.util.Collections;
 import java.util.HashSet;
 import java.util.LinkedHashMap;
-import java.util.LinkedList;
 import java.util.List;
 import java.util.Map;
 import java.util.Set;
@@ -23,7 +22,7 @@ import org.drools.planner.examples.ras2012.util.model.OccupationTracker.Builder;
 
 public class ArcProgression implements Directed {
 
-    private final LinkedList<Arc>            orderedArcs        = new LinkedList<Arc>();
+    private final List<Arc>                  orderedArcs        = new ArrayList<Arc>();
     private final SortedMap<BigDecimal, Arc> milestones         = new TreeMap<BigDecimal, Arc>();
     private final Map<Node, Arc>             arcsPerOrigin      = new LinkedHashMap<Node, Arc>();
     private final Map<Node, Arc>             arcsPerDestination = new LinkedHashMap<Node, Arc>();
@@ -45,8 +44,9 @@ public class ArcProgression implements Directed {
         while (arcs.size() != this.orderedArcs.size()) {
             for (final Arc a : arcs) {
                 if (a.getOrigin(directed) == startingNode) {
-                    this.orderedArcs.addLast(a);
-                    startingNode = this.orderedArcs.peekLast().getDestination(directed);
+                    this.orderedArcs.add(a);
+                    // find the node of the last arc in the progression
+                    startingNode = a.getDestination(directed);
                     break;
                 }
             }
@@ -134,7 +134,10 @@ public class ArcProgression implements Directed {
     }
 
     public Arc getDestination() {
-        return this.orderedArcs.peekLast();
+        if (this.isEmpty) {
+            throw new IllegalStateException("Empty progression has no destination.");
+        }
+        return this.orderedArcs.get(this.orderedArcs.size() - 1);
     }
 
     public BigDecimal getDistance(final Node end) {
@@ -251,7 +254,10 @@ public class ArcProgression implements Directed {
     }
 
     public Arc getOrigin() {
-        return this.orderedArcs.peekFirst();
+        if (this.isEmpty) {
+            throw new IllegalStateException("Empty progression has no origin.");
+        }
+        return this.orderedArcs.get(0);
     }
 
     public Arc getPrevious(final Arc a) {

### 2012-06-12 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/Route.java
Cover all corner cases when it comes to route possibility.

@@ -233,21 +233,26 @@ public class Route extends Visualizable implements Comparable<Route>, Directed {
         }
         // now traverse arcs to make sure every other condition is met
         for (final Arc a : this.progression.getArcs()) {
-            if (a.getTrack() != Track.SIDING) { // we only have rules for sidings
+            final boolean isSiding = a.getTrack() == Track.SIDING;
+            final boolean isSwitch = a.getTrack() == Track.SWITCH;
+            if (!isSwitch && !isSiding) { // we only have rules for sidings/switches
                 continue;
             }
             if (t.isHeavy()) {
                 /*
-                 * heavy trains must never use a siding when there is a meet-pass with another NSA train. this is the easiest
-                 * way to fulfill the requirement.
+                 * heavy trains must never use a siding/switch when there is a meet-pass with another NSA train. this is the
+                 * easiest way to fulfill the requirement.
                  */
                 return false;
             }
-            // hazmat trains disallowed to take sidings
+            // hazmat trains disallowed to take sidings/switches
             if (t.carriesHazardousMaterials()) {
                 return false;
             }
             // make sure the route doesn't contain a siding shorter than the train
+            if (isSwitch) {
+                continue;
+            }
             final int result = a.getLength().compareTo(t.getLength());
             if (result < 0) {
                 return false;

### 2012-06-12 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/Itinerary.java
Simplify code.

@@ -170,18 +170,13 @@ public final class Itinerary extends Visualizable {
             throw new IllegalArgumentException("Node cannot be null.");
         }
         final SortedMap<Long, Node> nodeEntryTimes = this.getSchedule();
-        long timeEntered = -1;
         for (final SortedMap.Entry<Long, Node> entry : nodeEntryTimes.entrySet()) {
             if (entry.getValue() == n) {
-                timeEntered = entry.getKey();
-                break;
+                return entry.getKey();
             }
         }
-        if (timeEntered < this.trainEntryTime) {
-            throw new IllegalStateException(
-                    "Proper node cannot be found! Possibly a bug in the algoritm.");
-        }
-        return timeEntered;
+        throw new IllegalStateException(
+                "Proper node cannot be found! Possibly a bug in the algoritm.");
     }
 
     public long getDelay() {

### 2012-06-12 ras2012-solver/src/test/java/org/drools/planner/examples/ras2012/ToyExampleSolutionTest.java
Visualize the example solution.

@@ -55,8 +55,12 @@ public class ToyExampleSolutionTest {
         throw new IllegalArgumentException("Train does not exist!");
     }
 
+    private static final File DIRECTORY = new File("data/tests/"
+                                                + ToyExampleSolutionTest.class.getName());
+
     @Before
     public void prepareSolution() {
+        DIRECTORY.mkdirs();
         // prepare B1 according to the example
         final Train b1 = this.locateTrain("B1");
         final ItineraryAssignment b1Assignment = this.SOLUTION.getAssignment(b1);
@@ -65,6 +69,7 @@ public class ToyExampleSolutionTest {
         long delay = 5316076 - this.getArrivalAtNode(b1Assignment.getItinerary(), Node.getNode(6));
         b1Assignment.getItinerary().setWaitTime(Node.getNode(9),
                 WaitTime.getWaitTime(delay, TimeUnit.MILLISECONDS));
+        b1Assignment.getRoute().visualize(new File(DIRECTORY, "B1.png"));
         // prepare C1 according to the example
         final Train c1 = this.locateTrain("C1");
         final ItineraryAssignment c1Assignment = this.SOLUTION.getAssignment(c1);
@@ -75,10 +80,14 @@ public class ToyExampleSolutionTest {
         delay = 8158664 - this.getArrivalAtNode(c1Assignment.getItinerary(), Node.getNode(6));
         c1Assignment.getItinerary().setWaitTime(Node.getNode(4),
                 WaitTime.getWaitTime(delay, TimeUnit.MILLISECONDS));
+        c1Assignment.getRoute().visualize(new File(DIRECTORY, "C1.png"));
         // prepare A1 according to the example
         final Train a1 = this.locateTrain("A1");
         final ItineraryAssignment a1Assignment = this.SOLUTION.getAssignment(a1);
         a1Assignment.setRoute(this.locateRoute(4));
+        a1Assignment.getRoute().visualize(new File(DIRECTORY, "A1.png"));
+        this.SOLUTION.visualize(new File(DIRECTORY, "solution.png"));
+        this.SOLUTION.getTerritory().visualize(new File(DIRECTORY, "territory.png"));
     }
 
     @Test

### 2012-06-12 ras2012-paper/src/main/latex/paper/paper.tex
Make the table smaller.
Updates to the paper.
A better solution.
Newly benchmarked results.
Slowly starting with the paper.

@@
 \usepackage{multirow}
+\usepackage{rotating}
 \author{Lukáš Petrovický}
@@
 \maketitle
-\listoffigures
-\listoftables
-\tableofcontents
 
-\section{Hello...}
+\begin{abstract}
+This article describes an entry to the 2012 RAS Problem Solving Competition, concerning dispatching on multi-track territories.
+The entry is based on the Drools Planner, a Java-based solver.
+On a reasonably recent computer, the resulting algorithm is able to produce feasible results within a minute and has been fine-tuned to provide best results in under 3 minutes.
+Source code to the entry is open source and well documented.
+\end{abstract}
 
-...world!
+\section{Introduction}
+
+\section{Describing the solution}
+
+\section{Implementation}
+
+\section{Conclusion}
 
@@
 
-\input{RDS2.txt-13829.tex}.
+\section{Resolved systems}
+
+In this section, we show the best solutions reached for each data set.
+All these were reached within 3 minutes in a single-threaded run, using Intel i7 Q820 processor running Fedora 17 and 1 GB of heap space inside Java 7 runtime environment.
+The specific solver configurations used to reach these solutions may not have been the same every time.
+
+Please note that via the benchmarking functionality of Drools planner (see above), users may be able to fine-tune the algorithm.
+This fine-tuning can be focused either on providing better solutions or on faster turnaround times.
+
+\include{RDS1.txt-1582}.
+
+\include{RDS2.txt-8948}.
+
+\include{RDS3.txt-11697}.
+
+\include{TOY.txt-808}.
+
+\section{Visualizations}
+
+\begin{figure}
+\centering
+\includegraphics[width=150mm,angle=90]{solution.png}
+\caption{RAS DATA SET TOY example territory. Undirected graph, arcs have descriptions.}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=150mm,angle=90]{B1.png}
+\centering
+\caption{RAS DATA SET TOY example, route of Train B1. Directed graph where green marks the origin, red the destination and blue is where the train is allowed to wait.}
+\end{figure}
+
+In this section, we show examples of visualizations that the solution is capable of providing.
+These visualizations have been rendered on the fly using the JUNG library\footnote{Java Universal Network/Graph framework, http://jung.sourceforge.net}.
 

### 2012-06-14 ras2012-benchmark/src/main/java/org/drools/planner/examples/ras2012/Benchmark.java
Benchmark configs.

@@ -15,12 +15,18 @@ public class Benchmark {
 
     @SuppressWarnings({ "rawtypes", "unchecked" })
     private static Map getTemplateData() {
-        int[] hardSA = new int[] { 0, 5, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100 };
-        int[] softSA = new int[] { 0, 5, 10, 50, 100, 500, 1000, 5000, 10000 };
+        int[] hardSA = new int[] { 10, 20, 30, 40, 50, 60, 70, 80, 90, 100 };
+        int[] softSA = new int[] { 10, 100, 1000 };
+        int[] selections = new int[] { 4, 8, 16 };
+        int[] tabus = new int[] { 100, 1000, 10000 };
         Collection<int[]> SAs = new LinkedList<int[]>();
         for (int hard : hardSA) {
             for (int soft : softSA) {
-                SAs.add(new int[] { hard, soft });
+                for (int selection : selections) {
+                    for (int tabu : tabus) {
+                        SAs.add(new int[] { hard, soft, selection, tabu });
+                    }
+                }
             }
         }
         Map result = new HashMap();

### 2012-06-16 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/util/SolutionIO.java
Implement the validation test.
OPEN - #21: Implement solution validator 
http://github.com/triceo/ras12/issues/issue/21

@@ -125,8 +125,18 @@ public class SolutionIO {
 
     private static Document parseXml(final File result) throws SAXException, IOException,
             ParserConfigurationException {
-        // read the document and strip unnecessary stuff
-        String content = SolutionIO.readFile(result);
+        final String content = SolutionIO.readFile(result);
+        return SolutionIO.parseXml(content);
+    }
+
+    private static Document parseXml(final InputStream result) throws IOException,
+            ParserConfigurationException, SAXException {
+        final String content = SolutionIO.readStream(result);
+        return SolutionIO.parseXml(content);
+    }
+
+    private static Document parseXml(String content) throws IOException,
+            ParserConfigurationException, SAXException {
         if (content.length() == 0) {
             throw new IOException("There was a problem reading the XML file.");
         }
@@ -264,6 +274,10 @@ public class SolutionIO {
         } catch (final FileNotFoundException e) {
             return "";
         }
+    }
+
+    private static String readStream(final InputStream is) {
+        return new Scanner(is).useDelimiter("\\A").next();
     }
 
     private static BigDecimal tokenToBigDecimal(final Token t) {
@@ -542,9 +556,7 @@ public class SolutionIO {
         InputStream is = null;
         try {
             is = new FileInputStream(inputSolutionFile);
-            final DataSetParser p = new DataSetParser(is);
-            p.parse();
-            return this.createSolution(p);
+            return this.read(is);
         } catch (final FileNotFoundException e) {
             throw new IllegalArgumentException("Solution file doesn't exist: " + inputSolutionFile,
                     e);
@@ -570,6 +582,22 @@ public class SolutionIO {
             throw new IllegalStateException("Problem reading XML file.", e);
         }
         return solution;
+    }
+
+    public ProblemSolution read(final InputStream inputSolution) throws ParseException {
+        final DataSetParser p = new DataSetParser(inputSolution);
+        p.parse();
+        return this.createSolution(p);
+    }
+
+    public ProblemSolution read(final InputStream inputSolution, final InputStream result) {
+        try {
+            final ProblemSolution solution = this.read(inputSolution);
+            SolutionIO.processXmlTrains(SolutionIO.parseXml(result), solution);
+            return solution;
+        } catch (final Exception e) {
+            throw new IllegalStateException("Problem reading XML file.", e);
+        }
     }
 
     @SuppressWarnings({ "rawtypes", "unchecked" })

### 2012-06-16 ras2012-solver/src/test/java/org/drools/planner/examples/ras2012/util/SolutionIOTest.java

@@ -1,0 +1,108 @@
+package org.drools.planner.examples.ras2012.util;
+
+import java.io.File;
+import java.io.InputStream;
+import java.util.Collection;
+import java.util.LinkedList;
+import java.util.List;
+
+import org.drools.planner.core.score.buildin.hardandsoft.DefaultHardAndSoftScore;
+import org.drools.planner.core.score.buildin.hardandsoft.HardAndSoftScore;
+import org.drools.planner.examples.ras2012.ProblemSolution;
+import org.drools.planner.examples.ras2012.ScoreCalculator;
+import org.junit.Assert;
+import org.junit.Test;
+import org.junit.runner.RunWith;
+import org.junit.runners.Parameterized;
+import org.junit.runners.Parameterized.Parameters;
+
+@RunWith(Parameterized.class)
+public class SolutionIOTest {
+
+    private static enum SolutionType {
+
+        RDS1("/org/drools/planner/examples/ras2012/parser/validParserInput1.txt"), RDS2(
+                "/org/drools/planner/examples/ras2012/parser/validParserInput2.txt"), RDS3(
+                "/org/drools/planner/examples/ras2012/parser/validParserInput3.txt");
+
+        private final String definition;
+
+        SolutionType(final String definition) {
+            this.definition = definition;
+        }
+
+        public InputStream getDefinition() {
+            return SolutionIOTest.class.getResourceAsStream(this.definition);
+        }
+    }
+
+    private static Object[] getResource(final SolutionType solutionType, final int hard,
+            final int soft) {
+        return new Object[] {
+                solutionType.getDefinition(),
+                SolutionIOTest.class.getResourceAsStream(solutionType.name() + ".txt" + soft
+                        + ".xml"), hard, soft };
+    }
+
+    @Parameters
+    public static Collection<Object[]> getResources() {
+        final List<Object[]> resources = new LinkedList<Object[]>();
+        resources.add(SolutionIOTest.getResource(SolutionType.RDS1, 0, -1280));
+        resources.add(SolutionIOTest.getResource(SolutionType.RDS1, 0, -1848));
+        resources.add(SolutionIOTest.getResource(SolutionType.RDS1, 0, -2107));
+        resources.add(SolutionIOTest.getResource(SolutionType.RDS1, 0, -2146));
+        resources.add(SolutionIOTest.getResource(SolutionType.RDS1, 0, -3030));
[…]

### 2012-06-16 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/util/SolutionIO.java
Destinations need different handling.
OPEN - #21: Implement solution validator 
http://github.com/triceo/ras12/issues/issue/21

@@ -213,16 +213,27 @@ public class SolutionIO {
             final boolean isMovement = e.getName().equals("movement");
             final boolean isDestination = e.getName().equals("destination");
             reachedDestination = reachedDestination || isDestination;
-            if (isMovement || isDestination) {
+            if (isMovement) {
                 movementCount++;
                 final long time = new BigDecimal(e.getAttributeValue("entry")).multiply(
                         BigDecimal.valueOf(1000)).longValue();
-                final Node n = isDestination ? train.getDestination() : SolutionIO.locateArc(
-                        solution, e.getAttribute("arc").getValue()).getOrigin(train);
+                final Node n = SolutionIO.locateArc(solution, e.getAttribute("arc").getValue())
+                        .getOrigin(train);
                 final long actualTime = i.getArrivalTime(n);
                 final long delay = time - actualTime;
                 if (delay != 0) {
                     i.setWaitTime(n, WaitTime.getWaitTime(delay, TimeUnit.MILLISECONDS));
+                }
+            } else if (isDestination) {
+                movementCount++;
+                final long time = new BigDecimal(e.getAttributeValue("entry")).multiply(
+                        BigDecimal.valueOf(1000)).longValue();
+                final Node n = train.getDestination();
+                final long actualTime = i.getArrivalTime(n);
+                final long delay = time - actualTime;
+                if (delay != 0) {
+                    i.setWaitTime(properRoute.getProgression().getPreviousNode(n),
+                            WaitTime.getWaitTime(delay, TimeUnit.MILLISECONDS));
                 }
             } else {
                 throw new IllegalStateException("Train " + train.getName()

### 2012-06-16 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/util/SolutionIO.java
Complete the validation test. Now fix the bugs.
OPEN - #21: Implement solution validator 
http://github.com/triceo/ras12/issues/issue/21

@@ -4,9 +4,12 @@ import java.io.ByteArrayInputStream;
 import java.io.File;
 import java.io.FileInputStream;
 import java.io.FileNotFoundException;
+import java.io.FileOutputStream;
 import java.io.FileWriter;
 import java.io.IOException;
 import java.io.InputStream;
+import java.io.OutputStream;
+import java.io.OutputStreamWriter;
 import java.math.BigDecimal;
 import java.util.ArrayList;
 import java.util.Collection;
@@ -287,6 +290,7 @@ public class SolutionIO {
         }
     }
 
+    @SuppressWarnings("resource")
     private static String readStream(final InputStream is) {
         return new Scanner(is).useDelimiter("\\A").next();
     }
@@ -626,19 +630,28 @@ public class SolutionIO {
         }
     }
 
-    @SuppressWarnings({ "rawtypes", "unchecked" })
     public void writeXML(final ProblemSolution solution, final File outputSolutionFile) {
+        try {
+            writeXML(solution, new FileOutputStream(outputSolutionFile));
+        } catch (FileNotFoundException e) {
+            SolutionIO.logger.error("Failed writing " + solution.getName() + " into "
+                    + outputSolutionFile, e);
+        }
+    }
+
+    @SuppressWarnings({ "rawtypes", "unchecked" })
+    public void writeXML(final ProblemSolution solution, final OutputStream outputSolution) {
         try {
             final Map map = new HashMap();
             map.put("trains", this.prepareXmlData(solution));
             map.put("name", solution.getName());
             this.freemarker.getTemplate("schedule.xml.ftl").process(map,
-                    new FileWriter(outputSolutionFile));
+                    new OutputStreamWriter(outputSolution));
         } catch (final TemplateException e) {
             SolutionIO.logger.error("Failed processing XML schedule template.", e);
         } catch (final IOException e) {
             SolutionIO.logger.error("Failed writing " + solution.getName() + " into "
-                    + outputSolutionFile, e);
+                    + outputSolution, e);
         }
     }
 }

### 2012-06-16 ras2012-solver/src/test/java/org/drools/planner/examples/ras2012/util/SolutionIOTest.java

@@ -1,10 +1,13 @@
 package org.drools.planner.examples.ras2012.util;
 
+import java.io.ByteArrayOutputStream;
 import java.io.File;
+import java.io.IOException;
 import java.io.InputStream;
 import java.util.Collection;
 import java.util.LinkedList;
 import java.util.List;
+import java.util.Scanner;
 
 import org.drools.planner.core.score.buildin.hardandsoft.DefaultHardAndSoftScore;
 import org.drools.planner.core.score.buildin.hardandsoft.HardAndSoftScore;
@@ -38,10 +41,8 @@ public class SolutionIOTest {
 
     private static Object[] getResource(final SolutionType solutionType, final int hard,
             final int soft) {
-        return new Object[] {
-                solutionType.getDefinition(),
-                SolutionIOTest.class.getResourceAsStream(solutionType.name() + ".txt" + soft
-                        + ".xml"), hard, soft };
+        return new Object[] { solutionType.getDefinition(),
+                solutionType.name() + ".txt" + soft + ".xml", hard, soft };
     }
 
     @Parameters
@@ -79,30 +80,43 @@ public class SolutionIOTest {
         return resources;
     }
 
-    private final InputStream      solution, definition;
+    private final InputStream      definition;
+    private final String           solutionResource;
     private final HardAndSoftScore score;
 
     private static final File      parent = new File("data/tests/"
                                                   + SolutionIOTest.class.getCanonicalName());
 
-    public SolutionIOTest(final InputStream solutionDefinition, final InputStream solution,
+    public SolutionIOTest(final InputStream solutionDefinition, final String solution,
             final int hardScore, final int softScore) {
         this.definition = solutionDefinition;
-        this.solution = solution;
+        this.solutionResource = solution;
         this.score = DefaultHardAndSoftScore.valueOf(hardScore, softScore);
     }
 
+    @SuppressWarnings("resource")
     @Test
-    public void testSolutionValidation() {
+    public void testSolutionValidation() throws IOException {
         final SolutionIO io = new SolutionIO();
-        final ProblemSolution result = io.read(this.definition, this.solution);
+        final ProblemSolution result = io.read(this.definition,
+                SolutionIOTest.class.getResourceAsStream(this.solutionResource));
         if (!SolutionIOTest.parent.exists()) {
             SolutionIOTest.parent.mkdirs();
         }
-        io.writeXML(result,
[…]

### 2012-06-16 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/Itinerary.java
More informative exception.

@@ -378,7 +378,8 @@ public final class Itinerary extends Visualizable {
 
     public WaitTime setWaitTime(final Node n, final WaitTime w) {
         if (!this.getRoute().getProgression().getWaitPoints().contains(n)) {
-            throw new IllegalArgumentException(n + " not a wait point: " + this);
+            throw new IllegalArgumentException(n + " not a wait point: " + this + ". Cannot set "
+                    + w + ".");
         }
         if (w == null) {
             return this.removeWaitTime(n);

### 2012-06-17 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/ScoreCalculator.java
OPEN - task 23: Remove old unused parts of API http://github.com/triceo/ras12/issues/issue/23

@@ -110,17 +110,6 @@ public class ScoreCalculator extends AbstractIncrementalScoreCalculator<ProblemS
         return DefaultHardAndSoftScore.valueOf(-conflicts, -penalty);
     }
 
-    private void clearEveryCache() {
-        this.wantTimePenalties.clear();
-        this.unpreferredTracksPenalties.clear();
-        this.scheduleAdherencePenalties.clear();
-        this.delayPenalties.clear();
-        this.conflicts = new ConflictRegistry(
-                (int) this.solution.getPlanningHorizon(TimeUnit.MINUTES)
-                        * ScoreCalculator.OCCUPATION_CHECKS_PER_MINUTE + 1);
-        this.entries = new EntryRegistry(Node.count());
-    }
-
     public int getDelayPenalty(final Itinerary i) {
         final long delay = i.getDelay(this.solution.getPlanningHorizon(TimeUnit.MILLISECONDS));
         if (delay <= 0) {
@@ -139,7 +128,7 @@ public class ScoreCalculator extends AbstractIncrementalScoreCalculator<ProblemS
             case SET_WAIT_TIME:
                 final Node previousToModifiedNode = i.getRoute().getProgression()
                         .getPreviousNode(lastChange.getRight());
-                if (previousToModifiedNode != null && i.isNodeOnRoute(previousToModifiedNode)) {
+                if (previousToModifiedNode != null && i.hasNode(previousToModifiedNode)) {
                     // start re-calculating occupied arcs from the first change in the itinerary
                     return i.getArrivalTime(previousToModifiedNode);
                 }
@@ -231,10 +220,10 @@ public class ScoreCalculator extends AbstractIncrementalScoreCalculator<ProblemS
         final Itinerary i = ia.getItinerary();
         this.entries.resetTimes(t);
         for (final Node n : ia.getRoute().getProgression().getNodes()) {
-            if (!i.isNodeOnRoute(n)) {
+            if (!i.hasNode(n)) {
                 continue;
             }
-            long leaveTime = i.getLeaveTime(n);
+            final long leaveTime = i.getLeaveTime(n);
             if (leaveTime <= this.solution.getPlanningHorizon(TimeUnit.MILLISECONDS)) {
                 this.entries.setTimes(n, t, i.getArrivalTime(n), i.getLeaveTime(n));
             }
@@ -270,7 +259,14 @@ public class ScoreCalculator extends AbstractIncrementalScoreCalculator<ProblemS
     @Override
[…]

### 2012-06-17 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/Arc.java

@@ -13,8 +13,6 @@ public class Arc extends Section {
 
     private final BigDecimal           lengthInMiles;
 
-    private final String               asString;
-
     public Arc(final Track t, final BigDecimal lengthInMiles, final Node westNode,
             final Node eastNode) {
         super(westNode, eastNode);
@@ -26,7 +24,6 @@ public class Arc extends Section {
         }
         this.track = t;
         this.lengthInMiles = lengthInMiles;
-        this.asString = this.toStringInternal();
     }
 
     @Override
@@ -66,10 +63,6 @@ public class Arc extends Section {
 
     @Override
     public String toString() {
-        return this.asString;
-    }
-
-    private String toStringInternal() {
         final StringBuilder builder = new StringBuilder();
         builder.append("Arc (").append(this.track.getSymbol()).append(") [lengthInMiles=")
                 .append(this.lengthInMiles).append(", section=").append(super.toString())

### 2012-06-17 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/Itinerary.java

@@ -5,6 +5,7 @@ import java.math.BigDecimal;
 import java.util.Collection;
 import java.util.Collections;
 import java.util.HashMap;
+import java.util.LinkedHashSet;
 import java.util.Map;
 import java.util.SortedMap;
 import java.util.TreeMap;
@@ -42,8 +43,7 @@ public final class Itinerary extends Visualizable {
     private final Train                        train;
 
     private final AtomicBoolean                scheduleCacheValid    = new AtomicBoolean(false);
-    private final Collection<Node>             nodesEnRoute;
-    private final Collection<Arc>              arcsEnRoute;
+    private final Collection<Node>             hasNodes              = new LinkedHashSet<Node>();
     private final SortedMap<Long, Node>        scheduleCache         = new TreeMap<Long, Node>();
     private final SortedMap<Long, Arc>         scheduleCacheWithArcs = new TreeMap<Long, Arc>();
     private final long                         trainEntryTime;
@@ -71,16 +71,21 @@ public final class Itinerary extends Visualizable {
         this.route = r;
         this.train = t;
         this.trainEntryTime = t.getEntryTime(Itinerary.DEFAULT_TIME_UNIT);
-        final ArcProgression enRoute = this.getRoute().getProgression()
-                .tail(this.getTrain().getOrigin());
-        this.nodesEnRoute = enRoute.getNodes();
-        this.arcsEnRoute = enRoute.getArcs();
+
+        final ArcProgression progression = this.getRoute().getProgression();
+        Node currentNode = this.getTrain().getOrigin();
+        do {
+            this.hasNodes.add(currentNode);
+        } while ((currentNode = progression.getNextNode(currentNode)) != this.getTrain()
+                .getDestination());
+        this.hasNodes.add(currentNode);
+
         // initialize the maintenance windows
         if (maintenanceWindows != null) {
             for (final MaintenanceWindow mow : maintenanceWindows) {
                 final Node origin = mow.getOrigin(t);
                 final Node destination = mow.getDestination(t);
-                if (this.isNodeOnRoute(origin) && this.isNodeOnRoute(destination)) {
+                if (this.hasNode(origin) && this.hasNode(destination)) {
                     this.maintenances.put(origin, mow);
                 }
             }
@@ -96,7 +101,8 @@ public final class Itinerary extends Visualizable {
         int i = 0;
         long previousTime = 0;
         Arc previousArc = null;
-        for (final Arc currentArc : this.arcsEnRoute) {
[…]

### 2012-06-17 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/Node.java

@@ -8,7 +8,6 @@ import org.apache.commons.lang3.builder.CompareToBuilder;
 public class Node implements Comparable<Node> {
 
     private final int                       id;
-    private final String                    asString;
 
     private static final Map<Integer, Node> nodes = new HashMap<Integer, Node>();
 
@@ -31,7 +30,6 @@ public class Node implements Comparable<Node> {
 
     private Node(final int id) {
         this.id = id;
-        this.asString = this.toStringInternal();
     }
 
     @Override
@@ -45,13 +43,7 @@ public class Node implements Comparable<Node> {
 
     @Override
     public String toString() {
-        return this.asString;
-    }
-
-    private String toStringInternal() {
-        final StringBuilder builder = new StringBuilder();
-        builder.append("Node #").append(this.id);
-        return builder.toString();
+        return "Node #" + this.getId();
     }
 
 }

### 2012-06-17 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/Route.java

@@ -186,7 +186,7 @@ public class Route extends Visualizable implements Comparable<Route>, Directed {
                 final BigDecimal length = a.getLength();
                 final int speed = this.isEastbound() ? a.getTrack().getSpeedEastbound() : a
                         .getTrack().getSpeedWestbound();
-                result += Converter.getTimeFromSpeedAndDistance(speed, length);
+                result += Converter.getTimeFromSpeedAndDistance(BigDecimal.valueOf(speed), length);
             }
             this.travellingTimeInMilliseconds = result;
         }

### 2012-06-17 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/move/WaitTimeAssignmentMove.java

@@ -124,7 +124,7 @@ public class WaitTimeAssignmentMove implements Move {
         if (this.assignment.getRoute() != this.route) {
             return false;
         }
-        if (this.assignment.getItinerary().isNodeOnRoute(this.node)) {
+        if (this.assignment.getItinerary().hasNode(this.node)) {
             return this.waitTime != this.previousWaitTime;
         } else {
             return false;

### 2012-06-17 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/util/Converter.java

@@ -27,8 +27,4 @@ public class Converter {
                 .longValue();
     }
 
-    public static long getTimeFromSpeedAndDistance(final int speedInMPH, final BigDecimal distance) {
-        return Converter.getTimeFromSpeedAndDistance(BigDecimal.valueOf(speedInMPH), distance);
-    }
-
 }

### 2012-06-17 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/util/SolutionIO.java

@@ -632,8 +632,8 @@ public class SolutionIO {
 
     public void writeXML(final ProblemSolution solution, final File outputSolutionFile) {
         try {
-            writeXML(solution, new FileOutputStream(outputSolutionFile));
-        } catch (FileNotFoundException e) {
+            this.writeXML(solution, new FileOutputStream(outputSolutionFile));
+        } catch (final FileNotFoundException e) {
             SolutionIO.logger.error("Failed writing " + solution.getName() + " into "
                     + outputSolutionFile, e);
         }

### 2012-06-17 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/util/model/ArcProgression.java

@@ -311,28 +311,6 @@ public class ArcProgression implements Directed {
         return this.arcsPerOrigin.get(n);
     }
 
-    /**
-     * Returns arc progression that is a subset of this arc progression. The subset starts at the beginning and ends with the
-     * arc that has the specified node as its ending.
-     * 
-     * @param n Node to end the arc progression.
-     * @return Subset from nodes 0 to n. If the node is the ending node, returns this. If the node is the starting node, returns
-     *         empty progression.
-     */
-    public ArcProgression head(final Node n) {
-        if (!this.nodes.contains(n)) {
-            throw new IllegalArgumentException(n + " not in progression!");
-        } else if (n == this.getDestination().getDestination(this)) {
-            return this;
-        } else if (n == this.getOrigin().getOrigin(this)) {
-            return new ArcProgression(this, new Arc[0]);
-        }
-        final Arc a = this.getWithDestinationNode(n);
-        assert a != null;
-        final int indexOf = this.orderedArcs.indexOf(a);
-        return new ArcProgression(this, this.orderedArcs.subList(0, indexOf + 1));
-    }
-
     @Override
     public boolean isEastbound() {
         return this.isEastbound;
@@ -345,28 +323,6 @@ public class ArcProgression implements Directed {
     @Override
     public boolean isWestbound() {
         return !this.isEastbound();
-    }
-
-    /**
-     * Returns arc progression that is a subset of this arc progression. The subset starts with the arc that has the specified
-     * node as its beginning, ends where this progression ends.
-     * 
-     * @param n Node to start the arc progression.
-     * @return Subset from nodes n to end. If the node is the ending node, returns empty progression. If the node is the
-     *         starting node, returns this.
-     */
-    public ArcProgression tail(final Node n) {
-        if (!this.nodes.contains(n)) {
-            throw new IllegalArgumentException(n + " not in progression!");
-        } else if (n == this.getDestination().getDestination(this)) {
-            return new ArcProgression(this, new Arc[0]);
-        } else if (n == this.getOrigin().getOrigin(this)) {
-            return this;
-        }
-        final Arc a = this.getWithOriginNode(n);
-        assert a != null;
-        final int indexOf = this.orderedArcs.indexOf(a);
-        return new ArcProgression(this, this.orderedArcs.subList(indexOf, this.orderedArcs.size()));
     }
 
     @Override

### 2012-06-17 ras2012-solver/src/test/java/org/drools/planner/examples/ras2012/ToyExampleSolutionTest.java

@@ -18,9 +18,12 @@ import org.junit.Test;
 
 public class ToyExampleSolutionTest {
 
-    private final ProblemSolution SOLUTION = new SolutionIO()
-                                                   .read(new File(
-                                                           "src/main/resources/org/drools/planner/examples/ras2012/TOY.txt"));
+    private final ProblemSolution SOLUTION  = new SolutionIO()
+                                                    .read(new File(
+                                                            "src/main/resources/org/drools/planner/examples/ras2012/TOY.txt"));
+
+    private static final File     DIRECTORY = new File("data/tests/"
+                                                    + ToyExampleSolutionTest.class.getName());
 
     private long getArrivalAtNode(final Itinerary i, final Node n) {
         for (final SortedMap.Entry<Long, Node> entry : i.getSchedule().entrySet()) {
@@ -55,12 +58,9 @@ public class ToyExampleSolutionTest {
         throw new IllegalArgumentException("Train does not exist!");
     }
 
-    private static final File DIRECTORY = new File("data/tests/"
-                                                + ToyExampleSolutionTest.class.getName());
-
     @Before
     public void prepareSolution() {
-        DIRECTORY.mkdirs();
+        ToyExampleSolutionTest.DIRECTORY.mkdirs();
         // prepare B1 according to the example
         final Train b1 = this.locateTrain("B1");
         final ItineraryAssignment b1Assignment = this.SOLUTION.getAssignment(b1);
@@ -69,7 +69,7 @@ public class ToyExampleSolutionTest {
         long delay = 5316076 - this.getArrivalAtNode(b1Assignment.getItinerary(), Node.getNode(6));
         b1Assignment.getItinerary().setWaitTime(Node.getNode(9),
                 WaitTime.getWaitTime(delay, TimeUnit.MILLISECONDS));
-        b1Assignment.getRoute().visualize(new File(DIRECTORY, "B1.png"));
+        b1Assignment.getRoute().visualize(new File(ToyExampleSolutionTest.DIRECTORY, "B1.png"));
         // prepare C1 according to the example
         final Train c1 = this.locateTrain("C1");
         final ItineraryAssignment c1Assignment = this.SOLUTION.getAssignment(c1);
@@ -80,14 +80,15 @@ public class ToyExampleSolutionTest {
[…]

### 2012-06-17 ras2012-solver/src/test/java/org/drools/planner/examples/ras2012/model/ItineraryAssignmentTest.java

@@ -17,7 +17,7 @@ public class ItineraryAssignmentTest extends AbstractItineraryProviderBasedTest 
         final Route r = i.getRoute();
         final Map<Node, MaintenanceWindow> result = new HashMap<Node, MaintenanceWindow>();
         for (final MaintenanceWindow mow : mows) {
-            if (i.isNodeOnRoute(mow.getOrigin(r)) && i.isNodeOnRoute(mow.getDestination(r))) {
+            if (i.hasNode(mow.getOrigin(r)) && i.hasNode(mow.getDestination(r))) {
                 result.put(mow.getOrigin(r), mow);
             }
         }

### 2012-06-17 ras2012-solver/src/test/java/org/drools/planner/examples/ras2012/model/ItineraryTest.java

@@ -14,7 +14,6 @@ import java.util.concurrent.TimeUnit;
 import org.drools.planner.examples.ras2012.ProblemSolution;
 import org.drools.planner.examples.ras2012.util.Converter;
 import org.junit.Assert;
-import org.junit.Ignore;
 import org.junit.Test;
 import org.junit.runners.Parameterized.Parameters;
 
@@ -150,7 +149,7 @@ public class ItineraryTest extends AbstractItineraryProviderBasedTest {
         while ((currentArc = r.getProgression().getNextArc(currentArc)) != null) {
             // account for possible maintenance windows
             final Node n = currentArc.getOrigin(r);
-            if (!this.itinerary.isNodeOnRoute(n)) { // sometimes a train doesn't start at the beginning of a route
+            if (!this.itinerary.hasNode(n)) { // sometimes a train doesn't start at the beginning of a route
                 continue;
             }
             if (this.itinerary.getMaintenances().containsKey(n)
@@ -217,12 +216,6 @@ public class ItineraryTest extends AbstractItineraryProviderBasedTest {
                         occupiedArcs);
             }
         }
-    }
-
-    @Ignore
-    @Test
-    public void testGetSchedule() {
-        Assert.fail("Not yet implemented"); // TODO
     }
 
     @Override

### 2012-06-17 ras2012-solver/src/test/java/org/drools/planner/examples/ras2012/util/SolutionIOTest.java

@@ -104,13 +104,13 @@ public class SolutionIOTest {
             SolutionIOTest.parent.mkdirs();
         }
         // compare XML results
-        ByteArrayOutputStream baos = new ByteArrayOutputStream();
+        final ByteArrayOutputStream baos = new ByteArrayOutputStream();
         io.writeXML(result, baos);
-        String oldContent = new Scanner(
+        final String oldContent = new Scanner(
                 SolutionIOTest.class.getResourceAsStream(this.solutionResource))
                 .useDelimiter("\\A").next();
-        String newContent = baos.toString();
-        File f = new File(SolutionIOTest.parent, result.getName() + this.score.getSoftScore()
+        final String newContent = baos.toString();
+        final File f = new File(SolutionIOTest.parent, result.getName() + this.score.getSoftScore()
                 + ".xml");
         io.writeXML(result, f);
         Assert.assertTrue("XML files do not match. Check " + f, newContent.equals(oldContent));

### 2012-06-17 ras2012-solver/src/test/java/org/drools/planner/examples/ras2012/util/model/ArcProgressionTest.java

@@ -52,21 +52,6 @@ public class ArcProgressionTest {
     }
 
     @Test
-    public void testCompleteHeadAndTail() {
-        final Arc a = new Arc(Track.MAIN_0, BigDecimal.ONE, Node.getNode(0), Node.getNode(1));
-        final Arc b = new Arc(Track.MAIN_0, BigDecimal.ONE, Node.getNode(1), Node.getNode(2));
-        final Arc c = new Arc(Track.MAIN_0, BigDecimal.ONE, Node.getNode(2), Node.getNode(3));
-        final List<Arc> arcs = Arrays.asList(new Arc[] { a, b, c });
-        final ArcProgression p = new ArcProgression(this.route, arcs);
-
-        // testing complete head/tail
-        Assert.assertSame("Head ends with last node; result should be full.", p,
-                p.head(p.getDestination().getDestination(p)));
-        Assert.assertSame("Tail starts with first node; result should be full.", p,
-                p.tail(p.getOrigin().getOrigin(p)));
-    }
-
-    @Test
     public void testConstructor() {
         final Arc a = new Arc(Track.MAIN_0, BigDecimal.ONE, Node.getNode(0), Node.getNode(1));
         final Arc b = new Arc(Track.MAIN_0, BigDecimal.ONE, Node.getNode(1), Node.getNode(2));
@@ -108,21 +93,6 @@ public class ArcProgressionTest {
                 .contains(arc));
         Assert.assertFalse("Collection should not contain the never-inserted arc.", r
                 .getProgression().contains(arc2));
-    }
-
-    @Test
-    public void testEmptyHeadAndTail() {
-        final Arc a = new Arc(Track.MAIN_0, BigDecimal.ONE, Node.getNode(0), Node.getNode(1));
-        final Arc b = new Arc(Track.MAIN_0, BigDecimal.ONE, Node.getNode(1), Node.getNode(2));
-        final Arc c = new Arc(Track.MAIN_0, BigDecimal.ONE, Node.getNode(2), Node.getNode(3));
-        final List<Arc> arcs = Arrays.asList(new Arc[] { a, b, c });
-        final ArcProgression p = new ArcProgression(this.route, arcs);
-
-        // testing empty head/tail
-        Assert.assertEquals("Head ends with first node; result should be empty.", 0,
-                p.head(p.getOrigin().getOrigin(p)).countArcs());
-        Assert.assertEquals("Tail starts with last node; result should be empty.", 0,
[…]

### 2012-06-17 ras2012-paper/src/main/latex/paper/paper.tex
Benchmarked solutions.
Re-benchmark after the bugfix.
Newly benchmarked results.
Re-evaluate solutions.

@@
 
+\section{Achieved results}
+
+Running the algorithm described above 20 times over each data set, we have compiled a set of results, see Table \ref{table:results}.
+All these were reached within 3 minutes in a single-threaded run, using Intel i7 Q820 processor running Fedora 17 and 2 GB of heap space inside Java 7 runtime environment.
+
+\begin{table}
+\caption{Solution performance per data set.}
+\begin{tabular}{c||c|c|c|c|c|c}
+\hline \hline
+    & Best    & Q1      & Q2      & Q3      & Worst   & Std. dev. \\ 
+\hline
+TOY & \$808   & \$960   & \$1210  & \$1210  & \$1210  & 164 \\ 
+\hline 
+1   & \$1435  & \$1695  & \$1930  & \$2856  & \$5612  & 1219 \\ 
+\hline 
+2   & \$8097  & \$10364 & \$14865 & \$16177 & \$22486 & 4088 \\ 
+\hline 
+3   & \$10617 & \$13247 & \$15367 & \$17307 & \$21384 & 2986 \\ 
+\end{tabular} 
+\label{table:results}
+\end{table}
+
+Please note that via the benchmarking functionality of Drools Planner (see above), users may be able to fine-tune the algorithm to be focused either on providing better solutions or on faster turnaround times.
+
+For statistics of the resolved systems, see tables \ref{table:TOY-808.tex}, \ref{table:RDS1-1435.tex}, \ref{table:RDS2-8097.tex} and \ref{table:RDS3-10617.tex} respectively.
+
 \section{Conclusion}
@@
 In this section, we show the best solutions reached for each data set.
-All these were reached within 3 minutes in a single-threaded run, using Intel i7 Q820 processor running Fedora 17 and 1 GB of heap space inside Java 7 runtime environment.
-The specific solver configurations used to reach these solutions may not have been the same every time.
 
-Please note that via the benchmarking functionality of Drools planner (see above), users may be able to fine-tune the algorithm.
-This fine-tuning can be focused either on providing better solutions or on faster turnaround times.
-
-\include{RDS1.txt-1582}.
-
-\include{RDS2.txt-8948}.
-
-\include{RDS3.txt-11697}.
-
-\include{TOY.txt-808}.
+\include{TOY-808}
+\include{RDS1-1435}
+\include{RDS2-8097}
+\include{RDS3-10617}
 

### 2012-06-17 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/RAS2012.java
Allow for plotting lookup results.

@@ -4,16 +4,23 @@ import java.io.File;
 import java.io.FileInputStream;
 import java.io.FileNotFoundException;
 import java.io.InputStream;
+import java.util.ArrayList;
+import java.util.HashMap;
 import java.util.HashSet;
+import java.util.List;
+import java.util.Map;
 import java.util.Set;
+import java.util.concurrent.Callable;
 import java.util.concurrent.ExecutorService;
 import java.util.concurrent.Executors;
+import java.util.concurrent.Future;
 import java.util.concurrent.TimeUnit;
 
 import org.drools.planner.config.XmlSolverFactory;
 import org.drools.planner.core.Solver;
 import org.drools.planner.core.score.buildin.hardandsoft.HardAndSoftScore;
 import org.drools.planner.examples.ras2012.CLI.ApplicationMode;
+import org.drools.planner.examples.ras2012.util.Chart;
 import org.drools.planner.examples.ras2012.util.SolutionIO;
 import org.slf4j.Logger;
 import org.slf4j.LoggerFactory;
@@ -35,7 +42,7 @@ public class RAS2012 {
     /**
      * Executes Drools planner on a particular data set.
      */
-    private static class SolverRunner implements Runnable {
+    private static class SolverRunner implements Callable<HardAndSoftScore> {
 
         private final InputStream dataset;
         private final String      name;
@@ -51,7 +58,7 @@ public class RAS2012 {
         }
 
         @Override
-        public void run() {
+        public HardAndSoftScore call() {
             RAS2012.logger.info(this.name + " solver starting...");
             final SolutionIO io = new SolutionIO();
             ProblemSolution sol;
@@ -59,7 +66,7 @@ public class RAS2012 {
                 sol = io.read(this.dataset);
             } catch (final Exception e) {
                 RAS2012.logger.error("Solver " + this.name + " finished unexpectedly. Cause: ", e);
-                return;
+                return null;
             }
             // and now start solving
             final XmlSolverFactory configurer = new XmlSolverFactory();
@@ -68,32 +75,36 @@ public class RAS2012 {
             solver.setPlanningProblem(sol);
             solver.solve();
             // output the solution
-            final File targetFolder = new File("data/solutions");
-            if (!targetFolder.exists()) {
-                targetFolder.mkdirs();
+            if (!RAS2012.resultDir.exists()) {
+                RAS2012.resultDir.mkdirs();
             }
             sol = (ProblemSolution) solver.getBestSolution();
[…]

### 2012-06-17 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/util/Chart.java

@@ -1,0 +1,53 @@
+package org.drools.planner.examples.ras2012.util;
+
+import java.awt.Color;
+import java.io.File;
+import java.io.IOException;
+import java.util.List;
+
+import org.jfree.chart.ChartUtilities;
+import org.jfree.chart.JFreeChart;
+import org.jfree.chart.axis.CategoryAxis;
+import org.jfree.chart.axis.NumberAxis;
+import org.jfree.chart.plot.CategoryPlot;
+import org.jfree.chart.renderer.category.BoxAndWhiskerRenderer;
+import org.jfree.data.statistics.BoxAndWhiskerCategoryDataset;
+import org.jfree.data.statistics.DefaultBoxAndWhiskerCategoryDataset;
+import org.slf4j.Logger;
+import org.slf4j.LoggerFactory;
+
+public class Chart {
+
+    private static final Logger                       logger  = LoggerFactory
+                                                                      .getLogger(Chart.class);
+
+    private final DefaultBoxAndWhiskerCategoryDataset dataset = new DefaultBoxAndWhiskerCategoryDataset();
+
+    public void addData(final List<Integer> data, final String datasetName) {
+        this.dataset.add(data, "", datasetName);
+    }
+
+    public BoxAndWhiskerCategoryDataset getDataset() {
+        return this.dataset;
+    }
+
+    public boolean plot(final File folder, final String filename) {
+        final CategoryAxis xAxis = new CategoryAxis("Data Set");
+        final NumberAxis yAxis = new NumberAxis("Cost");
+        yAxis.setAutoRangeIncludesZero(false);
+        final BoxAndWhiskerRenderer renderer = new BoxAndWhiskerRenderer();
+        renderer.setFillBox(false);
+        final CategoryPlot plot = new CategoryPlot(this.dataset, xAxis, yAxis, renderer);
+
+        final JFreeChart chart = new JFreeChart(plot);
+        chart.setBackgroundPaint(Color.WHITE);
+        chart.removeLegend();
+        try {
+            ChartUtilities.saveChartAsPNG(new File(folder, filename + ".png"), chart, 1024, 768);
+            return true;
+        } catch (final IOException e) {
+            logger.warn("Charting failed.", e);
+            return false;
+        }
+    }
+}

### 2012-06-17 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/util/SolutionIO.java

@@ -56,6 +56,7 @@ import org.jdom2.Element;
 import org.jdom2.filter.ElementFilter;
 import org.jdom2.input.DOMBuilder;
 import org.jdom2.util.IteratorIterable;
+import org.jfree.data.statistics.BoxAndWhiskerCategoryDataset;
 import org.slf4j.Logger;
 import org.slf4j.LoggerFactory;
 import org.xml.sax.SAXException;
@@ -612,6 +613,32 @@ public class SolutionIO {
             return solution;
         } catch (final Exception e) {
             throw new IllegalStateException("Problem reading XML file.", e);
+        }
+    }
+
+    public void writeChart(final BoxAndWhiskerCategoryDataset dataset, final File outputSolutionFile) {
+        final String rowId = "";
+        final Map<String, List<Object>> results = new HashMap<String, List<Object>>();
+        results.put("solutions", new ArrayList<Object>());
+        for (final Object key : dataset.getColumnKeys()) {
+            final String columnId = (String) key;
+            final Map<String, Object> values = new HashMap<String, Object>();
+            values.put("name", columnId);
+            values.put("best", dataset.getMinOutlier(rowId, columnId));
+            values.put("worst", dataset.getMaxOutlier(rowId, columnId));
+            values.put("q1", dataset.getQ1Value(rowId, columnId));
+            values.put("q2", dataset.getMedianValue(rowId, columnId));
+            values.put("q3", dataset.getQ3Value(rowId, columnId));
+            results.get("solutions").add(values);
+        }
+        try {
+            this.freemarker.getTemplate("stats.tex.ftl").process(results,
+                    new FileWriter(outputSolutionFile));
+        } catch (final TemplateException e) {
+            SolutionIO.logger.error("Failed processing solution statistics template.", e);
+        } catch (final IOException e) {
+            SolutionIO.logger.error(
+                    "Failed writing solution statistics into " + outputSolutionFile, e);
         }
     }
 

### 2012-06-17 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/RAS2012.java
Forgot to return the number of lookups back to a reasonable value.

@@ -163,7 +163,7 @@ public class RAS2012 {
         for (final String entry : streams) {
             RAS2012.logger.info("Starting lookup for the best solutions on " + entry + "...");
             scores.put(entry, new ArrayList<Future<HardAndSoftScore>>());
-            for (int i = 0; i < 1; i++) {
+            for (int i = 0; i < 20; i++) {
                 RAS2012.logger.info("Scheduled attempt #" + i + ".");
                 scores.get(entry).add(
                         RAS2012.executor.submit(new SolverRunner(RAS2012.class

### 2012-06-23 ras2012-paper/src/main/latex/paper/paper.tex
Update the resolved systems.

--- 2012-06-12 ras2012-paper/src/main/latex/paper/paper.tex
@@
 
+\section{Achieved results}
+
+Running the algorithm described above 20 times over each data set, we have compiled a set of results, see Table \ref{table:result}.
+All these were reached within 3 minutes in a single-threaded run, using Intel i7 Q820 processor running Fedora 17 and 2 GB of heap space inside Java 7 runtime environment.
+
+\include{stats}
+
+\begin{figure}
+\centering
+\includegraphics[width=120mm]{chart.png}
+\caption{Plotting the data for the various resolved systems.}
+\end{figure}
+
+Please note that via the benchmarking functionality of Drools Planner, users may be able to fine-tune the algorithm to be focused either on providing better solutions or on faster turnaround times.
+Drools Planner even allows for retrieving the intermediate results of the algorithm and also modify the problem while it's being solved, which makes it the ideal tool for real-time planning.
+
+For statistics of the resolved systems, see tables \ref{table:TOY-808.tex}, \ref{table:RDS1-996.tex}, \ref{table:RDS2-7597.tex} and \ref{table:RDS3-11023.tex} respectively.
+
 \section{Conclusion}
@@
 In this section, we show the best solutions reached for each data set.
-All these were reached within 3 minutes in a single-threaded run, using Intel i7 Q820 processor running Fedora 17 and 1 GB of heap space inside Java 7 runtime environment.
-The specific solver configurations used to reach these solutions may not have been the same every time.
 
-Please note that via the benchmarking functionality of Drools planner (see above), users may be able to fine-tune the algorithm.
-This fine-tuning can be focused either on providing better solutions or on faster turnaround times.
-
-\include{RDS1.txt-1582}.
-
-\include{RDS2.txt-8948}.
-
-\include{RDS3.txt-11697}.
-
-\include{TOY.txt-808}.
+\include{TOY-808}
+\include{RDS1-996}
+\include{RDS2-7597}
+\include{RDS3-11023}
 

### 2012-06-23 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/RAS2012.java
Don't use the TOY solution.

@@ -157,7 +157,6 @@ public class RAS2012 {
         streams.add("RDS1");
         streams.add("RDS2");
         streams.add("RDS3");
-        streams.add("TOY");
         // prepare futures
         final Map<String, List<Future<HardAndSoftScore>>> scores = new HashMap<String, List<Future<HardAndSoftScore>>>();
         for (final String entry : streams) {

### 2012-07-14 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/ProblemSolution.java
Share visualizers.

@@ -48,6 +48,7 @@ public class ProblemSolution extends Visualizable implements Solution<HardAndSof
     private HardAndSoftScore                      score;
 
     private long                                  horizon     = 0;
+    private final GraphVisualizer                 visualizer;
 
     /**
      * Create a clone of an existing solution. This clone will be exactly the same as the original solution, except for the
@@ -66,6 +67,7 @@ public class ProblemSolution extends Visualizable implements Solution<HardAndSof
         }
         this.horizon = problem.horizon;
         this.score = problem.getScore();
+        this.visualizer = problem.visualizer;
     }
 
     /**
@@ -93,6 +95,12 @@ public class ProblemSolution extends Visualizable implements Solution<HardAndSof
             ia.setRoute(this.getTerritory().getBestRoute(t));
             this.assignments.put(t, ia);
         }
+        // prepare a visualizer
+        final Collection<Arc> arcs = new HashSet<Arc>();
+        for (final ItineraryAssignment ia : this.getAssignments()) {
+            arcs.addAll(ia.getRoute().getProgression().getArcs());
+        }
+        this.visualizer = new GraphVisualizer(arcs);
     }
 
     @Override
@@ -212,11 +220,7 @@ public class ProblemSolution extends Visualizable implements Solution<HardAndSof
 
     @Override
     public boolean visualize(final File target) {
-        final Collection<Arc> arcs = new HashSet<Arc>();
-        for (final ItineraryAssignment ia : this.getAssignments()) {
-            arcs.addAll(ia.getRoute().getProgression().getArcs());
-        }
-        return this.visualize(new GraphVisualizer(arcs), target);
+        return this.visualize(this.visualizer, target);
     }
 
 }

### 2012-07-14 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/Route.java

@@ -150,9 +150,12 @@ public class Route extends Visualizable implements Comparable<Route>, Directed {
 
     private final Map<Train, Boolean> routePossibilitiesCache      = new HashMap<Train, Boolean>();
 
+    private final RouteVisualizer     visualizer;
+
     private Route(final int id, final Arc... e) {
         this.id = id;
         this.progression = new ArcProgression(this, e);
+        this.visualizer = new RouteVisualizer(this);
     }
 
     @Override
@@ -323,6 +326,6 @@ public class Route extends Visualizable implements Comparable<Route>, Directed {
 
     @Override
     public boolean visualize(final File target) {
-        return this.visualize(new RouteVisualizer(this), target);
+        return this.visualize(this.visualizer, target);
     }
 }

### 2012-07-14 ras2012-paper/src/main/latex/paper/paper.tex
Newly benchmarked results.

@@
 
-Running the algorithm described above 20 times over each data set, we have compiled a set of results, see Table \ref{table:result}.
+Running the algorithm described above 50 times over each data set, we have compiled a set of results, see Table \ref{table:result}.
 All these were reached within 3 minutes in a single-threaded run, using Intel i7 Q820 processor running Fedora 17 and 2 GB of heap space inside Java 7 runtime environment.
@@
 
-For statistics of the resolved systems, see tables \ref{table:TOY-808.tex}, \ref{table:RDS1-996.tex}, \ref{table:RDS2-7597.tex} and \ref{table:RDS3-11023.tex} respectively.
+For statistics of the resolved systems, see tables \ref{table:RASDATASET1}, \ref{table:RASDATASET2} and \ref{table:RASDATASET3} respectively.
 
@@
 
-\include{TOY-808}
-\include{RDS1-996}
-\include{RDS2-7597}
-\include{RDS3-11023}
+\include{RDS1-1199}
+\include{RDS2-7955}
+\include{RDS3-11306}
 

### 2012-07-14 ras2012-benchmark/src/main/java/org/drools/planner/examples/ras2012/Benchmark.java

@@ -15,16 +15,19 @@ public class Benchmark {
 
     @SuppressWarnings({ "rawtypes", "unchecked" })
     private static Map getTemplateData() {
-        int[] hardSA = new int[] { 10, 20, 30, 40, 50, 60, 70, 80, 90, 100 };
+        int[] hardSA = new int[] { 10, 50, 100, 500 };
         int[] softSA = new int[] { 10, 100, 1000 };
-        int[] selections = new int[] { 4, 8, 16 };
-        int[] tabus = new int[] { 100, 1000, 10000 };
+        int[] selections = new int[] { 4 };
+        int[] moveTabus = new int[] { 1, 7, 19 };
+        int[] solutionTabus = new int[] { 1, 100, 1000, 10000 };
         Collection<int[]> SAs = new LinkedList<int[]>();
         for (int hard : hardSA) {
             for (int soft : softSA) {
                 for (int selection : selections) {
-                    for (int tabu : tabus) {
-                        SAs.add(new int[] { hard, soft, selection, tabu });
+                    for (int moveTabu : moveTabus) {
+                        for (int solutionTabu : solutionTabus) {
+                            SAs.add(new int[] { hard, soft, selection, moveTabu, solutionTabu });
+                        }
                     }
                 }
             }

### 2012-07-14 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/App.java

@@ -180,7 +180,7 @@ public class App {
         for (final String entry : streams) {
             App.logger.info("Starting lookup for the best solutions on " + entry + "...");
             scores.put(entry, new ArrayList<Future<HardAndSoftScore>>());
-            for (int i = 0; i < 20; i++) {
+            for (int i = 0; i < 50; i++) {
                 App.logger.info("Scheduled attempt #" + i + ".");
                 scores.get(entry).add(
                         App.executor.submit(new SolverRunner(App.class.getResourceAsStream(entry

### 2012-07-27 ras2012-paper/src/main/latex/paper/paper.tex
Adding significant portions of text.

@@
-\documentclass[10pt,a4paper,draft]{article}
+\documentclass[10pt,a4paper,final]{article}
 \usepackage[utf8x]{inputenc}
@@
 On a reasonably recent computer, the resulting algorithm is able to produce feasible results within a minute and has been fine-tuned to provide best results in under 3 minutes.
-Source code to the entry is open source and well documented.
+Source code of the entry is open and well documented.
 \end{abstract}
@@
 
-\section{Describing the solution}
+\section{Key concepts}
+
+In this section, we will describe the various approaches we have taken to implement the rules and constrains required by the competition.
+These are the key concepts leveraged in the algorithm, whose implementation will be described in later sections.
+
+\begin{description}
+\item[Train routes] From each particular data set, a territory is built.
+This territory is then analyzed and every possible route through that territory is found.
+Each train then holds a set of routes it is allowed to take, from which a solver will attempt to pick the best one\footnote{The competition rules state that the algorithm cannot assign trains to sidings.
+A case could be made that our approach here violates this rule.
+The algorithm may, indeed, explicitly put a train on a route with a siding.
+However, as the solution improves, this situation will inevitably be reduced to only occurring during a meet-pass event.
+Strictly speaking, we don't violate the rule at all, since we don't assign trains to sidings~-- we assign routes to trains.}.
+A train is not allowed to take a particular route when any of the following conditions hold true:
+
+\begin{itemize}
+\item Train is required to pass through a node that is not on the route. Examples would include situations when the route doesn't include the node with a schedule adherence requirement attached.
+\item Length of the train is bigger than the length of any one siding on the route.
+\item There are sidings on the route of a train carrying hazardous materials.
+\item There are sidings on the route of a heavy train. This is the easiest way to fulfill the requirement that a heavy train must never use a siding when meet-passing another train.
+\end{itemize}
+
+\item[Train wait points] Our solution only allows trains to stop at specific points on the route, called ``wait points''\footnote{If we allowed the train to stop at any node on the route, we would be significantly extending the state space of the solution.
+We believe that it wouldn't be beneficial to the quality of the resolved system and have therefore made a decision to only stop the trains where they don't block main tracks.}.
+A train can be held when its lead engine enters a node for which any of the following conditions holds true:
+
+\begin{itemize}
+\item The node is the source terminal of the train. The train is not in the territory yet.
+\item The node is the starting node of a crossover arc, keeping the other main track available for whatever trains to pass.
+\item The node is the ending node of a siding arc. The train occupies the siding.
+\end{itemize}
+
+Trains are automatically stopped for maintenance windows. However, no additional delays are allowed in those nodes~-- the train starts to move immediately after the maintenance window is over.
+
+\item[Train wait times] In each of the wait points described above, the algorithm may incur one of a selection of delays.
+These delays are called ``wait times'' and range from 1 minute to $X$ minutes, where $X$ is the minimal possible delay that would cause the train to fall outside the planning horizon waiting\footnote{This is done to once again reduce the state space by eliminating wait times that aren't viable.
+Once a train falls out of the planning horizon, we no longer care about its further journey.}.
+The wait times aren't distributed evenly across the stated range~-- they get sparser as we progress towards bigger numbers in the range.
+That should allow for both fine-grained and coarse-grained planning.
+
+\item[Itinerary] An itinerary of a train consists of a route and 0 or more wait times at the wait points on that route.
+
+\item[Solution] A solution is a set of itineraries, one for each train.
+The best found solution becomes a resolved system.
+A solution differs from the other by either of the trains taking a different route, waiting at different points and/or for different times.
+
+\item[Conflicts] The itinerary of each train is reached separately, without accounting for the whereabouts of any other train in the territory.
+This inevitably leads to solutions where two trains leverage the same part of the track at the same time, thus effectively colliding.
+These events are called ``conflicts'' and are treated as hard constraints~-- no feasible solution will ever demonstrate any conflict\footnote{We would think it goes without saying.
+However, we've been getting that question over and over when describing the algorithm to friends and colleagues.
+Hence the clarification.}.
+The same goes for the mandatory time lapse between any two trains.
+
+\end{description}
 
@@
 
+In this section, we will describe the implementation of the algorithm, its various properties, strength and drawbacks.
+
+\subsection{Drools Planner}
+
+Drools Planner\footnote{The Drools project, http://www.jboss.org/drools/} is an open-source Java-based solver that currently implements various tabu search techniques and simulated annealing.
+It allows for declaratively stating your problem, using the following basic concepts:
+
+\begin{description}
+\item[Planning entity] The entity used for planning, in our case it is the itinerary for a particular train.
+
+\item[Planning variable] Each planning entity hold various variables.
+By changing their values, Planner can arrive at a better or worse solution.
+In our case, planning variables are the route for the train and wait times in wait points.
+
+\item[Planning values] Each variable has a list of values applicable at the given point in time.
+For example, the itinerary has a list of routes that are available for its train to take.
+Each of them can become a value of the planning variable, resulting in a change in solution quality.
+
+\item[Move] is a way to change one solution into another, to cross from one state to another in a state space.
+Usually a move consists of changing the value of a planning variable.
+In our case, we know the following moves:
+
+\item[Scoring function] is a means of evaluating the quality of the solution.
+In our case, it contains of two parts~-- a hard score, representing the number of conflicts (see above), and a soft score, representing the cost of the system.
+
+Scoring function in Planner can take two forms - first, you can write it in plain old Java.
+Alternatively, you can leverage Drools Expert, giving at your disposal the full power of a ReteOO-based inference engine.
+We use the former.
+
+\end{description}
+
+When all these are provided to Planner, along with a solver configuration where we specify  the types of algorithms to use\footnote{Such as simulated annealing, entity tabu, solution tabu etc.} and termination conditions, it will start traversing the state space by trying the moves, looking for the best solution.
+While solving, Planner makes no I/O operations, utilizing only the CPU and RAM~-- which means that any performance improvement to the scoring function, the moves or to Planner itself will directly affect the quality of your scores.
+
+One of the benefits of Planner is that it is a Java application.
+That means neither the developer of the solver nor its user is limited to any particular operating system or platform.
+For our development and testing, we've been using Linux, but we've seen the submission perform on Windows as well.
+
+At the moment, Planner is a single-threaded application.
+Using multi-core CPUs will yield no benefit performance-wise, although you can leverage them to run multiple solvers at once~-- resolving one data set per core.
+
+\subsection{Typical flow of the algorithm}
+
+\begin{enumerate}
+\item Start processing.
+\item Parse the data set, re-constructing the territory.
+\item Turn the data set into an initial solution. Each train is assigned the fastest possible route through the territory with no wait times. 
+\item Launch the solver with the initial solution. The solver then iterates:
+
+\begin{enumerate}
+\item Generate all possible moves for the solution.
+\item Based on the user-specified solver configuration, some or all of the moves are evaluated. This is where the simulated annealing and tabu searches come into play.
+\item Should a better solution be found, store it.
+\item Repeat.
+\end{enumerate}
+
+\item When a user-specified termination condition is reached, the solver ends and reports the best found solution. The termination condition in our case is 3 minutes of run time.
+\item We take the solution and convert it into the submission format. Optionally, we visualize and tabulate.
+\item End processing.
+\end{enumerate}
+
+\subsection{Additional features}
+
+During the development, debugging and tuning of the algorithm, we have developed various tools to make our work easier\footnote{As a side note, although technically not a ``feature,'' the algorithm is written in such a way that it should be very easy to adapt it for changing circumstances.
+For example, changing the territory to a directed graph by introducing one-way arcs should only take a couple minutes and shouldn't affect performance.}.
+Following is a list of them.
+
+\begin{description}
+
+\item[Data set parser] is generated on-the-fly from a JavaCC-based\footnote{JavaCC parser/scanner generator, http://javacc.java.net/} grammar and can be easily extended or reused in other applications.
+
+\item[Basic visualizations] of train routes, territories and solutions are possible. We draw graphs of these to help with the analysis of the resolved systems and debugging of the algorithm.
+
+\item[Solution validation] for reading the resolved system's XML, calculating its cost and also visualizing it.
+Can be used for comparing various competition submissions against each other using a single algorithm.
+
+\end{description}
+
+\subsection{Usage}
+
+\subsubsection{Compiling}
+
+We've relied on the industry-standard Maven\footnote{Apache Maven, http://maven.apache.org/} build tool for our dependency management and release needs.
+It is open-source and available for every major operating system.
+Our source code is tracked on Github\footnote{Our homepage, http://triceo.github.com/ras12/} using the Git SCM.
+
+Once you have downloaded the Git and Maven tools, download the source code by executing the command (see Figure \ref{figure:git}).
+This will create a \texttt{ras12} folder in your current working directory, containing a one-for-one copy of our entire code repository.
+After switching into it, simply type another command (see Figure \ref{figure:maven}).
+
+Maven starts downloading our dependencies, building our code and running our extensive tests.
+It may take a while and it should end in success.
+When it's done, switch to a directory called \texttt{ras2012-solver}.
+You will notice a new file called \texttt{target/ras2012-solver-1.0-jar-with-dependencies.jar}.
+You can use it to run the solver, as described in the following section.
+
+\begin{figure}
+\centering
+\texttt{git clone git://git@github.com:triceo/ras12.git}
+\caption{Obtaining the source code.}
+\label{figure:git}
+\end{figure}
+
+\begin{figure}
+\centering
+\texttt{mvn install}
+\caption{Building the binary distribution.}
+\label{figure:maven}
+\end{figure}
+
+\begin{figure}
+\centering
+\texttt{java -jar <JAR-FILE> -r -d <PATH-TO-DATA-SET>}
+\caption{Running the application in a resolver mode.}
+\label{figure:run-resolver}
+\end{figure}
+
+\begin{figure}
+\centering
+\texttt{java -jar <JAR-FILE> -r -d <PATH-TO-DATA-SET> -x <SEED>}
+\caption{Running the application in a resolver mode, specifying a random seed.}
+\label{figure:run-resolver-seed}
+\end{figure}
+
+\begin{figure}
+\centering
+\texttt{java -jar <JAR-FILE> -e -d <PATH-TO-DATA-SET> -s <PATH-TO-XML>}
+\caption{Running the application in an evaluation mode.}
+\label{figure:run-evaluation}
+\end{figure}
+
+\begin{figure}
+\centering
+\texttt{java -jar <JAR-FILE> -l}
+\caption{Running the application in lookup mode.}
+\label{figure:run-lookup}
+\end{figure}
+
+\subsubsection{Running}
+
+To start resolving a data set with a fresh random seed, see Figure \ref{figure:run-resolver}. To provide a specific random seed, see Figure \ref{figure:run-resolver-seed}.
+
+To calculate cost for a specific resolved system, see Figure \ref{figure:run-evaluation}.
+
+To verify how your solver configuration performs on the three provided data sets, see Figure \ref{figure:run-lookup}.
+It will run the algorithm many times with a fresh random seed each time and provide a summary of the results, much like in Table \ref{table:result} and Figure \ref{figure:plot}.
+
+Please note that in all the figures, \texttt{<JAR-FILE>} refers to the \texttt{.jar} file from the previous section.
+
 \section{Achieved results}
 
-Running the algorithm described above 50 times over each data set, we have compiled a set of results, see Table \ref{table:result}.
+Running the algorithm 50 times over each data set and using a fresh random seed every time, we have compiled a set of results, see Table \ref{table:result}.
 All these were reached within 3 minutes in a single-threaded run, using Intel i7 Q820 processor running Fedora 17 and 2 GB of heap space inside Java 7 runtime environment.
@@
 \caption{Plotting the data for the various resolved systems.}
+\label{figure:plot}
 \end{figure}
@@
 Please note that via the benchmarking functionality of Drools Planner, users may be able to fine-tune the algorithm to be focused either on providing better solutions or on faster turnaround times.
-Drools Planner even allows for retrieving the intermediate results of the algorithm and also modify the problem while it's being solved, which makes it the ideal tool for real-time planning.
+Drools Planner even allows for retrieving the intermediate results of the algorithm and modifying the problem while it's being solved, which makes it the ideal tool for real-time planning.
 
@@
 
+We are submitting a well-documented solution to the stated problem which handles data sets both large and small, is easy to understand and extend and doesn't limit the user to any particular computing platform.
+We allow for tabulating and visualizing resolved systems for easier analysis, be they ours or competitors'.
+
+The submission is configured to work reasonably well on all three provided data sets.
+Users in the real world are encouraged to benchmark various solver configurations to come up with the one that matches the scale of their particular problem.
+
 \appendix
+
+Following are appendixes where we provide statistics of the best available resolved systems and examples of their visualizations.
+The paper would be complete even without them, for which reason we believe they shouldn't be counted toward the 10-page limit.
 
@@
 \centering
-\includegraphics[width=150mm,angle=90]{solution.png}
+\includegraphics[width=190mm,angle=90]{solution.png}
 \caption{RAS DATA SET TOY example territory. Undirected graph, arcs have descriptions.}
@@
 \begin{figure}
-\includegraphics[width=150mm,angle=90]{B1.png}
+\includegraphics[width=190mm,angle=90]{B1.png}
 \centering

### 2012-07-27 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/Track.java
Change the notation on sidings/switches.

@@ -11,7 +11,7 @@ import java.util.Map;
  */
 public enum Track {
 
-    MAIN_0('='), MAIN_1('<'), MAIN_2('>'), SWITCH('S', false), SIDING('/', false), CROSSOVER('C',
+    MAIN_0('='), MAIN_1('<'), MAIN_2('>'), SWITCH('/', false), SIDING('S', false), CROSSOVER('C',
             false);
 
     private final boolean                    isMainTrack;

### 2012-07-27 ras2012-paper/src/main/latex/paper/paper.tex
Tidying in TEX files.

@@
 
-Following are appendixes where we provide statistics of the best available resolved systems and examples of their visualizations.
+Following are appendixes where we provide statistics of the best available resolved systems.
 The paper would be complete even without them, for which reason we believe they shouldn't be counted toward the 10-page limit.
@@
 
-\section{Visualizations}
-
-\begin{figure}
-\centering
-\includegraphics[width=190mm,angle=90]{solution.png}
-\caption{RAS DATA SET TOY example territory. Undirected graph, arcs have descriptions.}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=190mm,angle=90]{B1.png}
-\centering
-\caption{RAS DATA SET TOY example, route of Train B1. Directed graph where green marks the origin, red the destination and blue is where the train is allowed to wait.}
-\end{figure}
-
-In this section, we show examples of visualizations that the solution is capable of providing.
-These visualizations have been rendered on the fly using the JUNG library\footnote{Java Universal Network/Graph framework, http://jung.sourceforge.net}.
-
 \end{document}

### 2012-07-27 ras2012-paper/src/main/latex/paper/paper.tex
Adding the introductory bits.

@@
 \begin{abstract}
-This article describes an entry to the 2012 RAS Problem Solving Competition, concerning dispatching on multi-track territories.
+This article describes an entry to the \emph{2012 RAS Problem Solving Competition}, concerning dispatching on multi-track territories.
 The entry is based on the Drools Planner, a Java-based solver.
@@
 
+Working in the field of expert systems, tackling complicated problems can easily become one's hobby.
+But even in your spare time, you still want to find out how you compare to others.
+Having heard of the \emph{2012 RAS Problem Solving Competition} at that point, registering wasn't at all a tough decision.
+After spending many a night on it, this is what emerged.
+
+We first describe the core concepts behind our submission, following by notes on the actual technical implementation.
+We show how the application can be used for both its primary purpose and also additional ones.
+We discuss the achieved results and potential for future improvement.
+In the appendices, we provide statistics for the best resolved systems.
+
+Not being members of any of the Operations Research communities or even academia at all, we sincerely apologize to the reader of this paper for unknowingly not following any and all conventions for writing such documents.
+The seasoned researcher must be frightened by the complete lack of citations and sourcing~-- however, there is a simple explanation.
+We~used no literature whatsoever in our process, except for the occasional software documentation on-line.
+
 \section{Key concepts}
 
-In this section, we will describe the various approaches we have taken to implement the rules and constrains required by the competition.
+In this section, we describe the various approaches we have taken to implement the rules and constrains required by the competition.
 These are the key concepts leveraged in the algorithm, whose implementation will be described in later sections.
@@
 Drools Planner\footnote{The Drools project, http://www.jboss.org/drools/} is an open-source Java-based solver that currently implements various tabu search techniques and simulated annealing.
-It allows for declaratively stating your problem, using the following basic concepts:
+It allows for stating your problem in a declarative manner, using the following basic concepts:
 

### 2012-07-27 ras2012-paper/src/main/latex/paper/paper.tex
More updates to the text.

@@
 Working in the field of expert systems, tackling complicated problems can easily become one's hobby.
-But even in your spare time, you still want to find out how you compare to others.
+But even in one's spare time, one still wants to find out how one compares to others.
 Having heard of the \emph{2012 RAS Problem Solving Competition} at that point, registering wasn't at all a tough decision.
@@
 \begin{description}
+
+\item[Meet-pass events] The most important concept to grasp in our case is the meet-pass event, or more precisely the lack thereof.
+The route of every train is planned independently of every other train and the only relevant criteria is the overall system's efficiency expressed by its cost.\footnote{While, of course, still honoring every rule of the competition.} This inevitably leads to trains being sent onto a collision course.
+These situations are called ``conflicts'' and are treated as hard constraints~-- no feasible solution will ever demonstrate any conflict\footnote{We would think it goes without saying.
+However, we've been getting that question over and over when describing the algorithm to friends and colleagues.
+Hence the clarification.}.
+The same goes for the mandatory time lapse between any two trains.
+
 \item[Train routes] From each particular data set, a territory is built.
 This territory is then analyzed and every possible route through that territory is found.
-Each train then holds a set of routes it is allowed to take, from which a solver will attempt to pick the best one\footnote{The competition rules state that the algorithm cannot assign trains to sidings.
-A case could be made that our approach here violates this rule.
-The algorithm may, indeed, explicitly put a train on a route with a siding.
-However, as the solution improves, this situation will inevitably be reduced to only occurring during a meet-pass event.
+Each train holds a set of routes it is allowed to take, from which solver will attempt to pick the best one\footnote{The competition rules state that the algorithm cannot assign trains to sidings.
+A case could be made that our approach here violates this rule~-- the algorithm indeed may put a train on a route with a siding.
+However, as the solution improves, this situation will be reduced to only occurring during what would be seen as a meet-pass event.
 Strictly speaking, we don't violate the rule at all, since we don't assign trains to sidings~-- we assign routes to trains.}.
@@
 
-\item[Conflicts] The itinerary of each train is reached separately, without accounting for the whereabouts of any other train in the territory.
-This inevitably leads to solutions where two trains leverage the same part of the track at the same time, thus effectively colliding.
-These events are called ``conflicts'' and are treated as hard constraints~-- no feasible solution will ever demonstrate any conflict\footnote{We would think it goes without saying.
-However, we've been getting that question over and over when describing the algorithm to friends and colleagues.
-Hence the clarification.}.
-The same goes for the mandatory time lapse between any two trains.
-
 \end{description}
@@
 
-In this section, we will describe the implementation of the algorithm, its various properties, strength and drawbacks.
+In this section, we will describe the implementation of the algorithm, its various properties, strengths and drawbacks.
 
@@
 \begin{description}
-\item[Planning entity] The entity used for planning, in our case it is the itinerary for a particular train.
+\item[Planning entity] in our case is the itinerary for a particular train.
 
-\item[Planning variable] Each planning entity hold various variables.
-By changing their values, Planner can arrive at a better or worse solution.
-In our case, planning variables are the route for the train and wait times in wait points.
+\item[Planning variable] is a part of the planning entity.
+By changing variable's value, Planner can arrive at a better or worse solution.
+In our case, we have two planning variables: the route for the train and wait times in wait points.
 
-\item[Planning values] Each variable has a list of values applicable at the given point in time.
+\item[Planning values] are possible values to apply to a planning variable at any given time.
 For example, the itinerary has a list of routes that are available for its train to take.
@@
 
+\begin{itemize}
+\item A move to assign or remove a wait time for a specific wait point in the itinerary.
+\item A move to change a route the train will take, removing any and all wait times from the itinerary.
+\end{itemize}
+
 \item[Scoring function] is a means of evaluating the quality of the solution.
@@
 We use the former.
+
+Implementing an efficient scoring function without the help of an inference engine means writing a lot of caching and also figuring out a mechanism to only incrementally update those parts of the score where variables actually changed.
+Those features you get ``for free'' from a system such as Drools Expert; however, the rules of the scoring function ended up being so complex that it actually justified writing code in the imperative Java language.
 
@@
 \item Based on the user-specified solver configuration, some or all of the moves are evaluated. This is where the simulated annealing and tabu searches come into play.
-\item Should a better solution be found, store it.
+\item Should any of the moves result in a better solution, store it.
 \item Repeat.
@@
 
-\item When a user-specified termination condition is reached, the solver ends and reports the best found solution. The termination condition in our case is 3 minutes of run time.
+\item When a user-specified termination condition is reached, the solver ends and reports the best found solution. The termination condition in our case is achieving 3 minutes of run time.
 \item We take the solution and convert it into the submission format. Optionally, we visualize and tabulate.
@@
 \centering
-\texttt{java -jar <JAR-FILE> -r -d <PATH-TO-DATA-SET>}
-\caption{Running the application in a resolver mode.}
+\texttt{java -jar <JAR-FILE> -r -d <PATH-TO-DATA-SET> -x <SEED>}
+\caption{Running the application in a resolver mode, optionally specifying a random seed.}
 \label{figure:run-resolver}
-\end{figure}
-
-\begin{figure}
-\centering
-\texttt{java -jar <JAR-FILE> -r -d <PATH-TO-DATA-SET> -x <SEED>}
-\caption{Running the application in a resolver mode, specifying a random seed.}
-\label{figure:run-resolver-seed}
 \end{figure}
@@
 
-To start resolving a data set with a fresh random seed, see Figure \ref{figure:run-resolver}. To provide a specific random seed, see Figure \ref{figure:run-resolver-seed}.
-
+To start resolving a data set, see Figure \ref{figure:run-resolver}.
 To calculate cost for a specific resolved system, see Figure \ref{figure:run-evaluation}.
-
 To verify how your solver configuration performs on the three provided data sets, see Figure \ref{figure:run-lookup}.
@@
 
-\include{stats}
+As we can see, the differences between the best and the worst scores produced by the algorithm are fairly large~-- even more than 100 \% for the smallest data set.
+This randomness can be eliminated by providing a fixed seed.
+That may not be practical however, since various data sets behave differently with different seeds.
+Best practice could be to use the median (``Q2'') value as a guide for what to expect from the algorithm generally.
+
+Please note that via the benchmarking functionality of Drools Planner\footnote{We had to improve that functionality inside Planner to allow for parallel benchmarking.
+Otherwise our benchmarks would probably still be running today.}, users may be able to fine-tune the algorithm to be focused either on providing better solutions or on faster turnaround times.
+Drools Planner even allows for retrieving the intermediate results of the algorithm and modifying the problem while it's being solved, which makes it the ideal tool for real-time planning.
+
+For statistics of the resolved systems, see tables \ref{table:RASDATASET1}, \ref{table:RASDATASET2} and \ref{table:RASDATASET3} respectively.
 
@@
 
-Please note that via the benchmarking functionality of Drools Planner, users may be able to fine-tune the algorithm to be focused either on providing better solutions or on faster turnaround times.
-Drools Planner even allows for retrieving the intermediate results of the algorithm and modifying the problem while it's being solved, which makes it the ideal tool for real-time planning.
+\include{stats}
 
-For statistics of the resolved systems, see tables \ref{table:RASDATASET1}, \ref{table:RASDATASET2} and \ref{table:RASDATASET3} respectively.
+\section{The future}
+
+Drools Planner is an actively developed project, which means it brings new features all the time.
+Some of the recently added features include just-in-time move selectors.
+That feature, when introduced into the code base of our solution, could significantly reduce memory footprint of the application, leaving less memory for garbage collection and thus speeding up the solver.
+However, it wasn't possible to incorporate it into the submission while still meeting the competition deadline.
+
+Also, it is possible to refactor the application's code to allow for using Planner's generic move facilities.
+That would make the application more light-weight by removing significant portions of our own code for handling moves, while at the same time benefiting from various optimizations incorporated in Drools Planner.
+
+Although much care has been taken to optimize the inner workings of our submission to be as fast as possible, there still remain places where more optimization is at hand.
+Future work could focus on eliminating these places with the help of a code profiler.
 
@@
 The submission is configured to work reasonably well on all three provided data sets.
-Users in the real world are encouraged to benchmark various solver configurations to come up with the one that matches the scale of their particular problem.
+Users in the real world are encouraged to benchmark various solver configurations to come up with the one that best matches the scale of their particular problem.
 

### 2012-08-10 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/ScoreCalculator.java
Don't count useless sidings after the planning horizon.
Will need to research for best solutions again.

@@ -176,13 +176,17 @@ public class ScoreCalculator extends AbstractIncrementalScoreCalculator<ProblemS
     public int getPenaltyForNoMeetPassOnSidings(final Itinerary i) {
         int penalty = 0;
         for (final Arc a : i.getRoute().getProgression().getArcs()) {
+            final Node destination = a.getDestination(i.getTrain());
             if (a.getTrack() != Track.SIDING) {
                 continue;
             }
-            if (!i.hasNode(a.getOrigin(i.getTrain())) || i.hasNode(a.getDestination(i.getTrain()))) {
+            if (!i.hasNode(a.getOrigin(i.getTrain())) || i.hasNode(destination)) {
                 continue;
             }
-            if (i.getWaitTime(a.getDestination(i.getTrain())) != null) {
+            if (!this.isInPlanningHorizon(i.getArrivalTime(destination))) {
+                continue;
+            }
+            if (i.getWaitTime(destination) != null) {
                 penalty++;
             }
         }

### 2012-08-10 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/util/SolutionIO.java
Fix the tests.

@@ -247,7 +247,7 @@ public class SolutionIO {
         if (movementCount == 0) {
             /* we need so big delay that the train completely falls out of the planning horizon. */
             i.setWaitTime(train.getOrigin(), WaitTime.getWaitTime(
-                    solution.getPlanningHorizon(TimeUnit.MILLISECONDS), TimeUnit.MILLISECONDS));
+                    solution.getPlanningHorizon(TimeUnit.MILLISECONDS) + 1, TimeUnit.MILLISECONDS));
         } else if (!reachedDestination) {
             /*
              * train needs to get so big a delay at the first node out the horizon so that every other node falls right out too.

### 2012-08-10 ras2012-solver/src/test/java/org/drools/planner/examples/ras2012/util/SolutionIOTest.java

@@ -26,7 +26,8 @@ public class SolutionIOTest {
 
         RDS1("/org/drools/planner/examples/ras2012/parser/validParserInput1.txt"), RDS2(
                 "/org/drools/planner/examples/ras2012/parser/validParserInput2.txt"), RDS3(
-                "/org/drools/planner/examples/ras2012/parser/validParserInput3.txt");
+                "/org/drools/planner/examples/ras2012/parser/validParserInput3.txt"), TOY(
+                "/org/drools/planner/examples/ras2012/parser/validParserInput4.txt");
 
         private final String definition;
 
@@ -41,42 +42,20 @@ public class SolutionIOTest {
 
     private static Object[] getResource(final SolutionType solutionType, final int hard,
             final int soft) {
-        return new Object[] { solutionType.getDefinition(),
-                solutionType.name() + ".txt" + soft + ".xml", hard, soft };
+        return new Object[] { solutionType.getDefinition(), solutionType.name() + soft + ".xml",
+                hard, soft };
     }
 
     @Parameters
     public static Collection<Object[]> getResources() {
         final List<Object[]> resources = new LinkedList<Object[]>();
-        resources.add(SolutionIOTest.getResource(SolutionType.RDS1, 0, -1160));
-        resources.add(SolutionIOTest.getResource(SolutionType.RDS1, 0, -1459));
-        resources.add(SolutionIOTest.getResource(SolutionType.RDS1, 0, -1570));
-        resources.add(SolutionIOTest.getResource(SolutionType.RDS1, 0, -2271));
-        resources.add(SolutionIOTest.getResource(SolutionType.RDS1, 0, -2623));
-        resources.add(SolutionIOTest.getResource(SolutionType.RDS1, 0, -2673));
-        resources.add(SolutionIOTest.getResource(SolutionType.RDS1, 0, -2943));
-        resources.add(SolutionIOTest.getResource(SolutionType.RDS1, 0, -5253));
-        resources.add(SolutionIOTest.getResource(SolutionType.RDS1, 0, -5502));
-        resources.add(SolutionIOTest.getResource(SolutionType.RDS1, 0, -7124));
-        resources.add(SolutionIOTest.getResource(SolutionType.RDS2, 0, -8764));
-        resources.add(SolutionIOTest.getResource(SolutionType.RDS2, 0, -8822));
[…]

### 2012-08-10 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/ScoreCalculator.java
Fix bug in useless sidings calculations.
Will need re-benchmarking.

@@ -180,7 +180,7 @@ public class ScoreCalculator extends AbstractIncrementalScoreCalculator<ProblemS
             if (a.getTrack() != Track.SIDING) {
                 continue;
             }
-            if (!i.hasNode(a.getOrigin(i.getTrain())) || i.hasNode(destination)) {
+            if (!i.hasNode(a.getOrigin(i.getTrain())) || !i.hasNode(destination)) {
                 continue;
             }
             if (!this.isInPlanningHorizon(i.getArrivalTime(destination))) {

### 2012-08-12 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/Itinerary.java
Finally final benchmark.

@@ -449,12 +449,6 @@ public final class Itinerary extends Visualizable {
         if (!this.getRoute().getProgression().getWaitPoints().contains(n)) {
             throw new IllegalArgumentException(n + " not a wait point: " + this + ". Cannot set "
                     + w + ".");
-        } else if (n == this.getTrain().getOrigin()
-                && this.getTrain().getOrigin() != this.getRoute().getProgression().getOrigin()
-                        .getOrigin(this.getRoute())) {
-            // train starts in the middle of the network; don't let it stop there
-            throw new IllegalArgumentException(
-                    "Train can't immediately stop when it starts in the middle of the territory.");
         }
         if (w == null) {
             return this.removeWaitTime(n);

### 2012-08-12 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/ItineraryAssignment.java

@@ -132,11 +132,6 @@ public final class ItineraryAssignment implements Cloneable {
         for (final Node waitPoint : this.getRoute().getProgression().getWaitPoints()) {
             if (!this.getItinerary().hasNode(waitPoint)) {
                 continue;
-            } else if (waitPoint == this.getTrain().getOrigin()
-                    && waitPoint != this.getRoute().getProgression().getOrigin()
-                            .getOrigin(this.getTrain())) {
-                // don't allow the train to stop in cases where it would start somewhere in the middle of the territory
-                continue;
             }
             final WaitTime existingWaitTime = this.getItinerary().getWaitTime(waitPoint);
             final long currentArrival = this.getItinerary().getArrivalTime(waitPoint);

### 2012-08-12 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/move/RouteReassignmentMove.java

@@ -9,6 +9,7 @@ import org.apache.commons.lang3.builder.HashCodeBuilder;
 import org.drools.planner.core.move.Move;
 import org.drools.planner.core.score.director.ScoreDirector;
 import org.drools.planner.examples.ras2012.ProblemSolution;
+import org.drools.planner.examples.ras2012.model.Itinerary;
 import org.drools.planner.examples.ras2012.model.ItineraryAssignment;
 import org.drools.planner.examples.ras2012.model.Node;
 import org.drools.planner.examples.ras2012.model.Route;
@@ -45,6 +46,13 @@ public class RouteReassignmentMove implements Move {
     public void doMove(final ScoreDirector scoreDirector) {
         this.assignment = this.initializeMove(scoreDirector);
         this.assignment.setRoute(this.route);
+        final Itinerary i = this.assignment.getItinerary();
+        for (final Map.Entry<Node, WaitTime> entry : this.previousWaitTimes.entrySet()) {
+            Node n = entry.getKey();
+            if (i.hasNode(n) && this.route.getProgression().getWaitPoints().contains(n)) {
+                i.setWaitTime(n, entry.getValue());
+            }
+        }
         scoreDirector.afterEntityAdded(this.assignment);
     }
 

### 2012-08-14 ras2012-paper/src/main/latex/paper/paper.tex
Update the best solutions found.
Finally final benchmark.
Finalized version of the paper.

@@
 This article describes an entry to the \emph{2012 RAS Problem Solving Competition}, concerning dispatching on multi-track territories.
-The entry is based on the Drools Planner, a Java-based solver.
-On a reasonably recent computer, the resulting algorithm is able to produce feasible results within a minute and has been fine-tuned to provide best results in under 3 minutes.
+The entry is based on Drools Planner, a Java-based solver.
+On a reasonably recent computer, the resulting algorithm is able to produce feasible solutions within seconds and has been fine-tuned to provide best results in under 3 minutes.
 Source code of the entry is open and well documented.
@@
 We show how the application can be used for both its primary purpose and also additional ones.
-We discuss the achieved results and potential for future improvement.
-In the appendices, we provide statistics for the best resolved systems.
+We discuss the achieved results and in the appendices, we provide statistics for the best resolved systems.
 
 Not being members of any of the Operations Research communities or even academia at all, we sincerely apologize to the reader of this paper for unknowingly not following any and all conventions for writing such documents.
-The seasoned researcher must be frightened by the complete lack of citations and sourcing~-- however, there is a simple explanation.
-We~used no literature whatsoever in our process, except for the occasional software documentation on-line.
+The seasoned researcher must be frightened by the complete lack of citations and sourcing~-- however, there is a simple explanation: we used no literature whatsoever in our process, except for the occasional software documentation on-line.
+
+Throughout this text, we will use the following words, meaning:
+
+\begin{description}
+\item[Algorithm] is the solver-related code and configuration being submitted.
+\item[Application] consists of the algorithm plus other non-essential code, such as the visualizers.
+\item[Solution] is a resolved system. It can either be feasible, having all the hard constraints met, or unfeasible.
+\item[Submission] consists of this paper, the example solutions and the application. It is our final output for the purposes of the competition.
+\end{description}
 
@@
 
-In this section, we describe the various approaches we have taken to implement the rules and constrains required by the competition.
+In this section, we describe the various approaches we have taken to implement the rules and constraints required by the competition.
 These are the key concepts leveraged in the algorithm, whose implementation will be described in later sections.
@@
 The route of every train is planned independently of every other train and the only relevant criteria is the overall system's efficiency expressed by its cost.\footnote{While, of course, still honoring every rule of the competition.} This inevitably leads to trains being sent onto a collision course.
-These situations are called ``conflicts'' and are treated as hard constraints~-- no feasible solution will ever demonstrate any conflict\footnote{We would think it goes without saying.
-However, we've been getting that question over and over when describing the algorithm to friends and colleagues.
-Hence the clarification.}.
-The same goes for the mandatory time lapse between any two trains.
+These situations are eliminated by enforcing the five-minute separation rule at each arc's entry points.
 
@@
 \begin{itemize}
-\item Train is required to pass through a node that is not on the route. Examples would include situations when the route doesn't include the node with a schedule adherence requirement attached.
+\item Train is required to pass through a node that is not on the route.
+Examples would include situations when the selected route doesn't include the node where the train has a schedule adherence requirement set.
 \item Length of the train is bigger than the length of any one siding on the route.
-\item There are sidings on the route of a train carrying hazardous materials.
-\item There are sidings on the route of a heavy train. This is the easiest way to fulfill the requirement that a heavy train must never use a siding when meet-passing another train.
+\item There are sidings on the route of a train carrying hazardous materials. This is to fulfill the requirement that such train can never take a siding.
+\item There are sidings on the route of a heavy train. This is the easiest way to fulfill the requirement that a heavy train must never use a siding when meet-passing another NSA train.
 \end{itemize}
 
-\item[Train wait points] Our solution only allows trains to stop at specific points on the route, called ``wait points''\footnote{If we allowed the train to stop at any node on the route, we would be significantly extending the state space of the solution.
-We believe that it wouldn't be beneficial to the quality of the resolved system and have therefore made a decision to only stop the trains where they don't block main tracks.}.
+\item[Train wait points] Our solution only allows trains to stop at specific points on the route, called ``wait points''\footnote{If we allowed the train to stop at any node on the route, we would be significantly extending the state space.
+Benchmarking the algorithm showed that it wouldn't be beneficial to the quality of the solution.
+We have therefore made a decision to only stop the trains where they don't block main tracks.}.
 A train can be held when its lead engine enters a node for which any of the following conditions holds true:
@@
 The wait times aren't distributed evenly across the stated range~-- they get sparser as we progress towards bigger numbers in the range.
-That should allow for both fine-grained and coarse-grained planning.
+That allows for both fine-grained and coarse-grained planning, while still keeping the number of possibilities relatively small.
 
@@
 \item A move to assign or remove a wait time for a specific wait point in the itinerary.
-\item A move to change a route the train will take, removing any and all wait times from the itinerary.
+\item A move to change a route the train will take. We keep the existing wait times, except obviously when their respective wait point isn't part of the new route.
 \end{itemize}
@@
 \item[Scoring function] is a means of evaluating the quality of the solution.
-In our case, it contains of two parts~-- a hard score, representing the number of conflicts (see above), and a soft score, representing the cost of the system.
+In our case, it consists of two parts, the hard and the soft score.
+The soft score represents the cost of the system as defined by the competition.
+The hard score is a sum of five-minute rule violations and the number of trains taking sidings where they do not wait.
+In a feasible solution, the hard score will be zero, meaning that each train obeys the five-minute rule and never takes a siding unless it also stops there.
 
@@
 Alternatively, you can leverage Drools Expert, giving at your disposal the full power of a ReteOO-based inference engine.
-We use the former.
-
-Implementing an efficient scoring function without the help of an inference engine means writing a lot of caching and also figuring out a mechanism to only incrementally update those parts of the score where variables actually changed.
-Those features you get ``for free'' from a system such as Drools Expert; however, the rules of the scoring function ended up being so complex that it actually justified writing code in the imperative Java language.
+We use the former\footnote{Implementing an efficient scoring function without the help of an inference engine means writing a lot of caching and also figuring out a mechanism to only incrementally update those parts of the score where variables actually changed.
+Those features you get ``for free'' from a system such as Drools Expert; however, the rules of the scoring function ended up being so complex that it actually justified writing code in the imperative Java language.}.
 
@@
 
-When all these are provided to Planner, along with a solver configuration where we specify  the types of algorithms to use\footnote{Such as simulated annealing, entity tabu, solution tabu etc.} and termination conditions, it will start traversing the state space by trying the moves, looking for the best solution.
-While solving, Planner makes no I/O operations, utilizing only the CPU and RAM~-- which means that any performance improvement to the scoring function, the moves or to Planner itself will directly affect the quality of your scores.
+When all these are provided to Planner, along with a solver configuration where we specify  the types of algorithms to use and termination conditions\footnote{Our algorithm is configured to terminate after it ran for 3 minutes and didn't improve the solution in the last 1000 steps.}, it will start traversing the state space by trying the moves, looking for the best solution.
+While solving, Planner makes no I/O operations, utilizing only the CPU and RAM~-- which means that any performance improvement to the scoring function, to the moves or to Planner itself will directly affect the quality of your scores.
 
@@
 
+At the time of submission, our application relies on a yet unreleased version of Drools Planner, directly from the master branch.
+Thus, we recommend using the compiled binary version of our application provided as part of our submission, since that is guaranteed to work.
+In case the reader wishes to compile the code for themselves, they may first have to resolve various compile-time or runtime problems arising from possible incompatibilities between the current Planner version and the future version being used.
+
 \subsubsection{Running}
@@
 
-Please note that in all the figures, \texttt{<JAR-FILE>} refers to the \texttt{.jar} file from the previous section.
+Please note that in all the figures, \texttt{<JAR-FILE>} refers to the \texttt{.jar} file from the previous section, or the one received as part of this submission.
 
@@
 
-\section{The future}
-
-Drools Planner is an actively developed project, which means it brings new features all the time.
-Some of the recently added features include just-in-time move selectors.
-That feature, when introduced into the code base of our solution, could significantly reduce memory footprint of the application, leaving less memory for garbage collection and thus speeding up the solver.
-However, it wasn't possible to incorporate it into the submission while still meeting the competition deadline.
-
-Also, it is possible to refactor the application's code to allow for using Planner's generic move facilities.
-That would make the application more light-weight by removing significant portions of our own code for handling moves, while at the same time benefiting from various optimizations incorporated in Drools Planner.
-
-Although much care has been taken to optimize the inner workings of our submission to be as fast as possible, there still remain places where more optimization is at hand.
-Future work could focus on eliminating these places with the help of a code profiler.
-
 \section{Conclusion}
 
-We are submitting a well-documented solution to the stated problem which handles data sets both large and small, is easy to understand and extend and doesn't limit the user to any particular computing platform.
+We are submitting a solution to the stated problem that is easy to understand and extend and doesn't limit the user to any particular computing platform.
 We allow for tabulating and visualizing resolved systems for easier analysis, be they ours or competitors'.
 
-The submission is configured to work reasonably well on all three provided data sets.
-Users in the real world are encouraged to benchmark various solver configurations to come up with the one that best matches the scale of their particular problem.
+On all the provided data sets, the algorithm reaches feasible solutions within seconds.
+We provide resolved systems that were reached within 3 minutes of starting the algorithm.
+However, the algorithm can be easily adapted to run continuously and provide results on-demand.
+
+The application is configured to work reasonably well for the three real-world data sets\footnote{Every plot and table in this paper has been generated by the algorithm on-the-fly.}.
+Users are encouraged to benchmark their own solver configurations based on their problem sizes, in order to maximize the algorithm's performance for their particular use case.
+
+API of the algorithm is well-documented and contains an extensive test suite. The output of both, as well as the source code, has been published at the submission's website.
 
@@
 
-\include{RDS1-1199}
-\include{RDS2-7955}
-\include{RDS3-11306}
+\include{RDS1-1829}
+\include{RDS2-8255}
+\include{RDS3-10777}
+\include{TOY-821}
 

### 2012-08-14 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/Itinerary.java
Don't enforce wait point checks in the Itinerary.

@@ -438,18 +438,13 @@ public final class Itinerary extends Visualizable {
     }
 
     /**
-     * Make the {@link Train} stop at a particular {@link Node}. Will cause wait time change (see
-     * {@link #getLatestWaitTimeChange()}).
+     * Make the {@link Train} stop at a particular {@link Node}.
      * 
      * @param n Node to stop the train at.
      * @param w How long to wait there.
      * @return The previous wait time or null if none.
      */
     public WaitTime setWaitTime(final Node n, final WaitTime w) {
-        if (!this.getRoute().getProgression().getWaitPoints().contains(n)) {
-            throw new IllegalArgumentException(n + " not a wait point: " + this + ". Cannot set "
-                    + w + ".");
-        }
         if (w == null) {
             return this.removeWaitTime(n);
         }

### 2012-08-14 ras2012-solver/src/test/java/org/drools/planner/examples/ras2012/model/SyntheticItineraryTest.java
Fix tests.

@@ -106,14 +106,6 @@ public class SyntheticItineraryTest {
     }
 
     @Test(expected = IllegalArgumentException.class)
-    public void testGetterAndSetterOutsideWaitPoint() {
-        final Itinerary i = new Itinerary(SyntheticItineraryTest.ROUTE_POSSIBLE,
-                SyntheticItineraryTest.TRAIN1);
-        final WaitTime wt = WaitTime.getWaitTime(10);
-        i.setWaitTime(SyntheticItineraryTest.END, wt);
-    }
-
-    @Test(expected = IllegalArgumentException.class)
     public void testImpossibleRoute() {
         new Itinerary(SyntheticItineraryTest.ROUTE_IMPOSSIBLE, SyntheticItineraryTest.TRAIN1);
     }

### 2012-08-14 ras2012-solver/src/test/java/org/drools/planner/examples/ras2012/util/SolutionIOTest.java

@@ -49,14 +49,12 @@ public class SolutionIOTest {
     @Parameters
     public static Collection<Object[]> getResources() {
         final List<Object[]> resources = new LinkedList<Object[]>();
-        resources.add(SolutionIOTest.getResource(SolutionType.TOY, 0, -1074));
-        resources.add(SolutionIOTest.getResource(SolutionType.TOY, 0, -1099));
-        resources.add(SolutionIOTest.getResource(SolutionType.RDS1, 0, -1981));
-        resources.add(SolutionIOTest.getResource(SolutionType.RDS1, 0, -6213));
-        resources.add(SolutionIOTest.getResource(SolutionType.RDS2, 0, -7249));
-        resources.add(SolutionIOTest.getResource(SolutionType.RDS2, 0, -18655));
-        resources.add(SolutionIOTest.getResource(SolutionType.RDS3, 0, -10033));
-        resources.add(SolutionIOTest.getResource(SolutionType.RDS3, 0, -31269));
+        resources.add(SolutionIOTest.getResource(SolutionType.RDS1, 0, -1829));
+        resources.add(SolutionIOTest.getResource(SolutionType.RDS1, 0, -5355));
+        resources.add(SolutionIOTest.getResource(SolutionType.RDS2, 0, -8255));
+        resources.add(SolutionIOTest.getResource(SolutionType.RDS2, 0, -13728));
+        resources.add(SolutionIOTest.getResource(SolutionType.RDS3, 0, -11063));
+        resources.add(SolutionIOTest.getResource(SolutionType.RDS3, 0, -16108));
         return resources;
     }
 

### 2012-08-14 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/ScoreCalculator.java
More javadocs.

@@ -41,6 +41,11 @@ import org.drools.planner.examples.ras2012.util.EntryRegistry;
  * </p>
  * 
  * <p>
+ * Also a part of hard score is the number of unused sidings. Every time a train takes a siding where it doesn't stop, it is
+ * penalized.
+ * </p>
+ * 
+ * <p>
  * Soft constraints, those that only affect score quality and not its feasibility, are all defined by the problem. They are:
  * </p>
  * 
@@ -173,6 +178,12 @@ public class ScoreCalculator extends AbstractIncrementalScoreCalculator<ProblemS
                 .intValue();
     }
 
+    /**
+     * Find the number of sidings where the train does not wait.
+     * 
+     * @param i The schedule in question.
+     * @return The number in question.
+     */
     public int getPenaltyForNoMeetPassOnSidings(final Itinerary i) {
         int penalty = 0;
         for (final Arc a : i.getRoute().getProgression().getArcs()) {

### 2012-08-14 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/Itinerary.java

@@ -407,8 +407,7 @@ public final class Itinerary extends Visualizable {
     }
 
     /**
-     * Don't stop the {@link Train} at a particular {@link Node} any more. Will cause wait time change (see
-     * {@link #getLatestWaitTimeChange()}).
+     * Don't stop the {@link Train} at a particular {@link Node} any more.
      * 
      * @param n Node in question.
      * @return The time it originally waited there.
@@ -426,8 +425,7 @@ public final class Itinerary extends Visualizable {
     }
 
     /**
-     * Don't stop the {@link Train} anywhere, except for {@link MaintenanceWindow}s. Will cause wait time change (see
-     * {@link #getLatestWaitTimeChange()}).
+     * Don't stop the {@link Train} anywhere, except for {@link MaintenanceWindow}s.
      */
     public void removeWaitTimes() {
         if (this.nodeWaitTimes.size() > 0) {

### 2012-08-14 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/ItineraryAssignment.java

@@ -122,6 +122,12 @@ public final class ItineraryAssignment implements Cloneable {
         return this.train;
     }
 
+    /**
+     * Calculates the possible wait times for every wait point on the route. Will be used by {@link WaitTimeAssignmentFactory}
+     * to assemble all the valid moves. See {@link #getWaitTimeAssignments()} for the rules by which a wait time is generated.
+     * 
+     * @return A collection of all the valid wait points and wait times.
+     */
     public Collection<WaitTimeAssignment> getWaitTimeAssignments() {
         if (this.route == null) {
             return Collections.emptyList();

### 2012-08-14 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/model/WaitTimeAssignment.java

@@ -1,5 +1,8 @@
 package org.drools.planner.examples.ras2012.model;
 
+/**
+ * Represents a relationship between a wait point and a wait time.
+ */
 public class WaitTimeAssignment {
 
     private final Node     node;

### 2012-08-14 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/move/RouteReassignmentMove.java

@@ -18,6 +18,10 @@ import org.drools.planner.examples.ras2012.model.WaitTime;
 import org.slf4j.Logger;
 import org.slf4j.LoggerFactory;
 
+/**
+ * This move causes a train to move on one route instead of the other. It will keep whatever existing wait times, if possible.
+ * 
+ */
 public class RouteReassignmentMove implements Move {
 
     private static final Logger logger = LoggerFactory.getLogger(RouteReassignmentMove.class);

### 2012-08-14 ras2012-solver/src/main/java/org/drools/planner/examples/ras2012/move/WaitTimeAssignmentMove.java

@@ -17,6 +17,9 @@ import org.drools.planner.examples.ras2012.model.WaitTime;
 import org.slf4j.Logger;
 import org.slf4j.LoggerFactory;
 
+/**
+ * This move will make a train stop in a wait point for a defined amount of time.
+ */
 public class WaitTimeAssignmentMove implements Move {
 
     private static final Logger logger = LoggerFactory.getLogger(WaitTimeAssignmentMove.class);
